### 2024-09-04 index.Rmd
treees

---
title: "Lab journal"
author: "by: Vero"
bibliography: references.bib

---

```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
# install.packages("remotes")
# remotes::install_github("rlesur/klippy") # is what makes it possible to hover over top right of code on website to COPY AND PASTE it 
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

----

<!--- The ---- is a visual line (page break looking mf) on the 
website 
Always ADD A SPACE BETWEEN A PAE BREAK AND the next comment, otherwise it will appear on page --->

<!--- This is the best way to write a comment. 
It also demonstrates, you can add html directly in your rmarkdown file. See: https://www.w3schools.com/html/html_comments.asp --->

<!--- to use a multi column page we use Pandoc's fenced Div. 
see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html 
But please note, we are not here to make a beautiful website, our intention is just to have a quick lab journal up and running --->

:::: {.columns}
::: {.column width="50%"}

# My Lab Journal
Last compiled on `r format(Sys.time(), '%B, %Y')`

*Describe lab journal here.*

<br>

----

## Open Science

Everyone can fork this repository. Use this example to journal your lab work and assignments.

We, Jochem Tolsma and Bas Hofstra, use this repository for our Social Network Analysis For Social Scientists course [@SNASS]. <!--- this is a reference, stored in references.bib. It will be automatically added at the bottom of the page --->

treeeeees

:::

::: {.column width="50%"}

![](images/cover.jpg)

:::
::::

## How to use

1. Fork the repository for this website and follow instructions on `README` to get started [here](https://github.com/JochemTolsma/labjournal).

2. Journal your work, by editing the .Rmd-files

3. ...

4. Change everything to make it your own!

<br>

Use the top menu to navigate through the journal.

----

# References