### 2024-09-11 RQdraft1.Rmd
week1push

---
title: "RQdraft1"
author: "Vero"
date: "2024-09-10"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```
# Brainstorm
- Journal prestige 
- Gender homphily 
- Age 
- repeated publications 
- nr of publications
- Dutch or English text publications
- gender segregation and field of publication (journal field)

---

# Research Questions
RQ1a: 
  To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Computing Sciences at Radboud University?

RQ1b:
  To what extent do the two departments differ in their gender homophily?

RQ2:
  To what extent does gender homophily in co-publishing influence the prestige of the publication journal?

RQ3:
  To what extent do structural network effects influence the gender homophily of each department?

- (1a+b needed? Would like more concise, not ask uneccessary questions)
- (2&3 seem random and disconnected; Journal prestige of interest tho, how to best include?)
- (necessity to have RQ3 that asks about structural networks effects?)

## Horta et al. (2022)
- male researchers preferred to co-publish with other male researchers, 
- and female researchers preferred to collaborate with other females.
- researchers of approximately similar ages tended to collaborate more with each other.
- geographical, cultural, & institutional proximity were strongest predictors of collaboration.
- analysis shows that in the social sciences, ACQUIRED attributes (particularly preferences regarding research agendas) take precedence over ASCRIBED attributes (e.g. gender or age)

- Horta et al. (2022) - RQs:
  - (1) What homophily preferences determine research collaborations among higher education researchers? 
  - (2) What are the most powerful homophily indicators in research collaborations in higher education research?

## Kwiek & Roszka (2021)
- Most male scientists collaborate solely with males.
- Most female scientists, in contrast, do not collaborate with females at all.
- All-female authorship marginal, all-male authorship pervasive, in all age groups.
- Articles written in mixed-sex collaboration published in more prestigious journals.

---

# References:

Horta, H., Feng, S., & Santos, J. M. (2022). Homophily in higher education research: a perspective based on co-authorships. Scientometrics, 127(1), 523–543. https://doi.org/10.1007/s11192-021-04227-z

Kwiek, M., & Roszka, W. (2021). Gender-based homophily in research: A large-scale study of man-woman collaboration. Journal of Informetrics, 15(3), 101171. https://doi.org/10.1016/j.joi.2021.101171

-