### 2024-09-11 _week1.Rmd

---
title: "_SNASS tutorial"
author: "TreesDude"
date: "2024-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r token}
#install.packages("gitcreds")
# gitcreds::gitcreds_set()
```

```{r}
# TO MAKENEW CHUNKS: ctrl + Alt + i

```

```{r}
########################### Title: Introducation to R for SNA Author: J Tolsma version: 30-10-2019

# start with clean workspace
rm(list = ls())

# install.packages I will need later here
install.packages("installr")  # you  first install packages
install.packages("foreign")
require(foreign)  # used to read in spss data files
require(tidyverse)

# define workdirectory, note the double backslashes
setwd("C:\Users\Vero\OneDrive - Radboud Universiteit\Documents\GitHub\labjournal")  # change to your own workdirectory
```

```{r}
# ignore the warnings ?read.spss note that I have saved the data files in a folder called
# 'addfiles'.

cv08 <- foreign::read.spss("addfiles\\Cultural_Changes_2008.sav", use.value.labels = T, to.data.frame = T)
cv10 <- foreign::read.spss("addfiles\\Cultural_Changes_2010.sav", use.value.labels = T, to.data.frame = T)

# normally I think setting use.value.labels=F is more convenient. Thus lets load the data again but
# now without labels
cv08_nolab <- foreign::read.spss("addfiles\\Cultural_Changes_2008.sav", use.value.labels = F, to.data.frame = T)
cv10_nolab <- foreign::read.spss("addfiles\\Cultural_Changes_2010.sav", use.value.labels = F, to.data.frame = T)

# finally, import the data using haven
cv08_haven <- haven::read_spss("addfiles\\Cultural_Changes_2008.sav")
cv10_haven <- haven::read_spss("addfiles\\Cultural_Changes_2010.sav")
```