### 2024-09-11 testjt.Rmd

---
title: "t62al"
author: "by: Vero"
bibliography: references.bib

---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

<!--- 
HOW TO MAKE TOKENS:
install.packages("gitcreds") 
gitcreds::gitcreds_set() --->

<!--- TO MAKE NEW CHUNKS: ctrl + Alt + i --->

```{r, eval=FALSE}
########################### Title: Introduction to R for SNA Author: J Tolsma version: 30-10-2019

# start with clean workspace
rm(list = ls())

# install.packages I will need later here
install.packages("installr")  # you  first install packages
install.packages("foreign")
install.packages("tidyverse")
```

```{r}
require(foreign)  # used to read in spss data files
require(tidyverse)

# define workdirectory, note the double backslashes
# setwd("C:\Users\Vero\OneDrive - Radboud Universiteit\Documents\GitHub\labjournal")  # change to your own workdirectory
```
```{r Cheat sheet}
###########Cheat Sheet##########

# variable labels 
# c("primary", "secondary", "tertiary")

# Operator symbols: 
# $
# == logical operator. What are the other logical operators in R?
# [] is used to subset elements from an object (e.g. dataframe/vector/matrix)
#  NA is used in R to define missing values. It means Not Applicable.

# Functions:
# summary(), attributes(), attr(), table(), names(), head(), fix(), View()

# as.character() 
# as.numeric()

# in tidy: 
  #the piping operator 
  # %>%, 
  # the left is transported to the (first argument of the) function on the right. Thus, in the example below, you see that in the second and third call to mutate I don’t have to tell the function which dataset I am using.

# mean(), 
# rowMeans(), 
# rowSums, 
# cbind(), 
# is.na(), 
# ifelse().

# logicals FALSE and TRUE can be summed? (FALSE equals 0 and TRUE equals 1).
```

```{r}
# ignore the warnings ?read.spss note that I have saved the data files in a folder called
# 'addfiles'.

cv08 <- foreign::read.spss("tutorial_data\\Cultural_Changes_2008.sav", use.value.labels = T, to.data.frame = T)
cv10 <- foreign::read.spss("tutorial_data\\Cultural_Changes_2010.sav", use.value.labels = T, to.data.frame = T)

# normally I think setting use.value.labels=F is more convenient. Thus lets load the data again but
# now WITHOUT LABELS 
cv08_nolab <- foreign::read.spss("tutorial_data\\Cultural_Changes_2008.sav", use.value.labels = F, to.data.frame = T)
cv10_nolab <- foreign::read.spss("tutorial_data\\Cultural_Changes_2010.sav", use.value.labels = F, to.data.frame = T)

# finally, import the data using haven
cv08_haven <- haven::read_spss("tutorial_data\\Cultural_Changes_2008.sav")
cv10_haven <- haven::read_spss("tutorial_data\\Cultural_Changes_2010.sav")

str(cv08) # str = asks for structure of an object 
str(cv08_nolab)
str(cv08_haven)

# acsessing variable:
# datasetname$variablename

str(cv08$lftop)  # a factor
str(cv08_nolab$lftop)  # a numeric variable
str(cv08_haven$lftop)  # a 'dbl+lbl' this stands for doubles, or real numbers, which are labeled  
# next to the data itself, attributes are stored
attributes(cv08$lftop)
attributes(cv08_nolab$lftop)
attributes(cv08_haven$lftop)

# to access specific attributes
attr(cv08_haven$lftop, "labels")

summary(cv08$lftop)
summary(cv08_nolab$lftop)
summary(cv08_haven$lftop)
table(cv08_haven$lftop, useNA = "always")

#LOOKING AT DATA 
names(cv08_haven)
summary(cv08_haven)
head(cv08_haven)

# fix(cv08_haven) #will produce an error
# fix(cv08)
# View(cv08_haven)

```

----
```{r A.7.1 Define Missings}

str(cv08_haven$lftop)
summary(cv08_haven$lftop)
attr(cv08_haven$lftop, "labels")
table(cv08_haven$lftop, useNA = "always")

# copy the original variable in a new one, and attach it to the dataset. 
cv08$lftop_new <- cv08$lftop

# replace those values of our new variable
cv08$lftop_new[cv08$lftop_new == "Onbekend"] <- NA

#Check if recode works 
table(cv08$lftop_new, useNA = "always")

levels(cv08$lftop_new)

# as numeric variable not as a factor (categorical).
str(cv08$lftop_new)

cv08$agen <- as.numeric(as.character(cv08$lftop_new))  
# how clumsy. we first convert the factor to a string and then to a numeric variable.
table(cv08$agen, useNA = "always")

str(cv08$agen)

```

```{r A.7.2 Tidy }

# install.packages("tidyverse")
library(tidyverse)

# ‘mutate’ the original dataset by adding a variable.
cv08_haven <- mutate(cv08_haven, lftop_new = lftop)

#Replace missings.Note: 99 is the “onbekend” category.
cv08_haven$lftop_new <- na_if(cv08_haven$lftop_new, 99)

#Normally: combine these two steps into one
cv08_haven <- mutate(cv08_haven, lftop_new = na_if(lftop, 99))

#check
table(cv08_haven$lftop_new, useNA = "always")

```
----
```{r A.8 Recode variables}

# var006n = second variable lets us recode the variable education (in CV08)
#create a new variable educ3 with three levels:
#1. primary
#2. secondary
#3. tertiary

#levels
levels(cv08$var006n)

# table 
table(cv08$var006n, useNA = "always")

# lets make it a numeric variable first
cv08$educn <- as.numeric(cv08$var006n)
# check
table(cv08$educn, useNA = "always")

# start with an empty variable
cv08$educ3 <- NA
# fill category by category
cv08$educ3[cv08$educn == 2 | cv08$educn == 3] <- 1
cv08$educ3[cv08$educn > 3 & cv08$educn < 8] <- 2
cv08$educ3[cv08$educn > 7 & cv08$educn < 11] <- 3

# check
table(cv08$educ3, useNA = "always")

prop.table(table(cv08$educ3, useNA = "always"))

# now educ3 is a numeric variable, we want it as factor
cv08$educ3 <- as.factor(cv08$educ3)
table(cv08$educ3, useNA = "always")

levels(cv08$educ3) <- c("primary", "secondary", "tertiary")
table(cv08$educ3, useNA = "always")
```

```{r A.8.2 Tidy}

# dplyr::recode   #doesnt work on labelled variables that were imported via haven package

# install.packages('labelled')
require(labelled)  # to be able to use the recode function on haven labelled variables

# inspect variable
str(cv08_haven$var006n)

attr(cv08_haven$var006n, "labels")

table(cv08_haven$var006n, useNA = "always")

# recode values, all missings as one value
cv08_haven <- mutate(cv08_haven, educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2,
    `4` = 2, `5` = 2, `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE)

# replace missing values with NA.
cv08_haven <- mutate(cv08_haven, educ3 = na_if(educ3, -9))

# make educ3 a factor
cv08_haven <- mutate(cv08_haven, educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary",
    "tertiary")))

# check
table(cv08_haven$educ3, useNA = "always")

#the piping operator 
  # %>%, 
  # the left is transported to the (first argument of the) function on the right. Thus, in the example below, you see that in the second and third call to mutate I don’t have to tell the function which dataset I am using.
# in Base R (since 4.1.0) operator: |>

cv08_haven <- mutate(cv08_haven, educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2,
    `4` = 2, `5` = 2, `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE) %>%
    mutate(educ3 = na_if(educ3, -9)) %>%
    mutate(educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

# even more tidy:
cv08_haven <- cv08_haven %>%
    mutate(educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2,
        `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE) %>%
    mutate(educ3 = na_if(educ3, -9)) %>%
    mutate(educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

# the MOST tidy:
cv08_haven <- cv08_haven %>%
    mutate(educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2,
        `6` = 3, `7` = 3, `8` = 3, `10` = -9, .keep_value_labels = FALSE), educ3 = na_if(educ3, -9),
        educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))
```
----
```{r A.9 Means and counting specific values}
# A.9.1. R Base

# Step 1: Look at the various 
summary(cv08$int055)
summary(cv08$int056)
summary(cv08$int057)

# Step 2: Make numeric 
cv08$int055n <- as.numeric(cv08$int055)
table(cv08$int055n, useNA = "always")

cv08$int056n <- as.numeric(cv08$int056)
table(cv08$int056n, useNA = "always")

cv08$int057n <- as.numeric(cv08$int057)
table(cv08$int057n, useNA = "always")

# Step 3: Define missings and recode
cv08$int055n[cv08$int055n < 5] <- NA
cv08$int055n <- cv08$int055n - 4

cv08$int056n[cv08$int056n < 5] <- NA
cv08$int056n <- cv08$int056n - 4

cv08$int057n[cv08$int057n < 5] <- NA
cv08$int057n <- cv08$int057n - 4

# Step 4: Calculate means
# How does the function mean work in R?
mean(cv08$int055n)  # whoops
mean(cv08$int055n, na.rm = TRUE)  # works.  but not what we want. 
# na.rm -> tells function whether or not to remove NA values from the calculation

mean(c(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)  # works but not what we want.

# calculate a mean for each row/respondent. 
testmeans <- rowMeans(cbind(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)
head(testmeans)  # yes!
# head() function : Return the First or Last Part of an Object

# What we really want is a mean but only if there is a maximum of 1 NA in the three variables.
# lets first count how many missings we have for each respondent
nmis <- rowSums(is.na(cbind(cv08$int055n, cv08$int056n, cv08$int057n)))

# ?is.na ?rowSums
testmeans <- ifelse(nmis < 2, testmeans, NA)

# add the calculated means to our dataset
cv08$int_mean <- testmeans

```

----

----

# References