### 2024-09-16 Webscraping-Chapter8-tutorial.Rmd

---
title: "Chapter 8 - SNASS"
author: "Verooo"
date: "2024-09-16"
output: html_document
---
#########################################Title: Webscraping in R Author: Bas Hofstra Version:
######################################### 29-07-2

# start with clean workspace
rm(list = ls())

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
#install.packages('httr') # we don't need this for now require(httr)
install.packages("xml2")
require(xml2)
install.packages("rvest")
require(rvest)
install.packages("devtools")
require(devtools)
#Note we're doing something different here. We're installing a *latest* version directly from
#GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

require(scholar)

#define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'