### 2024-09-20 RQdraft3.Rmd
editsheadingw3

---
title: "RQ draft 3"
author: "Verooo"
date: "2024-09-20"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Questions

**RQ1a:**

-   To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Political Science Department at Radboud University?

**RQ1b:**

-   To what extent do the two departments differ in their gender homophily?

**RQ2:**

-   How to does gender homophily affect collaboration differently for collaborations between staff who have published in low vs. high prestige journals?

# Overview:

-   Sample:

    -   Radboud Sociology Dptm.

    -   Radboud Pol Sci Dptm.

-   Time frame: 2019 - 2023 ( Why? Well... because. )

-   Data:

    -   names

    -   publications

        -   connected to authors (for collaborations)

    -   initiator: position in list ofcollaborators

    -   sex/gender (f,m/w,m)

        -   using name data bank

    -   journals of the publications

        -   prestige: their ?-Index

        -   --\> categorize ? scale?

-   age

    -   time since first publication

        -   **until last publication? until present?**

# Questions

-   How scrape position of name in list of collaborators?

-

# To-Do

-   Read up on G-index, H-index, prestige indicators of journals

    -   decide which one as indicator

-