### 2024-10-02 week5.Rmd

---
title: "Week 5"
author: "Verooo"
date: "2024-10-02"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes while discussing other peeps plots
Using RSiena for evolutin of network 
Wave, Ego alters 
RSciena: Ego, alter, wave 
Reorganize into an array for RSciena

Set size of isolates to SMALL 
put all isolates in one place 
then you have enough room left to have a more "zoomed in" frame for the actual collaborator networks

Or just remove all isolates lol 
```{r}

```