### 2024-10-09 week6.Rmd
week6page

---
title: "Week 6"
author: "Verooo"
date: "2024-10-09"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

::: {.column width="50%"} ![](images/krill-yourself.jpg)

# "Kudos make you run!" - paper discussion

Franken, R., Bekhuis, H., & Tolsma, J. (2022). Kudos make you run! How runners influence each other on the online social network Strava. Social Networks, 72, 151–164. <https://doi.org/10.1016/j.socnet.2022.10.001>

## Event vs. relation

-   Giving a like (aka. a "kudos" on the Strava platform) is an *event* but RSiena looks at *relationships,* which are more constant.

```{r}

```