### 2024-10-27 week6.Rmd
journalupdatew6bug

---
title: "Week 6"
author: "Verooo"
date: "2024-10-09"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

::: {.column width="50%"} 
![](images/krill-yourself.jpg)

# "Kudos make you run!" - paper discussion

Franken, R., Bekhuis, H., & Tolsma, J. (2022). Kudos make you run! How runners influence each other on the online social network Strava. Social Networks, 72, 151–164. <https://doi.org/10.1016/j.socnet.2022.10.001>

## Tie meaning

### Event vs. relation

-   Giving a like (aka. a "kudos" on the Strava platform) is an *event* but RSiena looks at *relationships,* which are more constant.

## Demarkation of sample

important to define

## what we wanna know

behavior of egos as consequence of their network *ties*

## Reason for why RSiena

in paper: closest to causality

not for this paper but mabe fits to in our projects in this course: to control for higher macro level influences

## Main network statistic

### Frequency, volume, (and intensity) of doing sport

-   Frequency (how often you do sports, aka. nr. of activities)

-   Volume (how long),

-   (and intensity) of doing sport

    -   intensity is unethical to scrape so it was not scraped (heart rate, pace etc.)

## theory of influence

if validated by peers (you have more kudos) more likely to run again

If you have higher in-degree, leads to behavioral increase

ACTUALITY: the more **different** people you received kudos by in a month: the more likely they were to do activity again

The more out degrees the more people tended to act in similar way

people giving kudos to people with lower activity more likely to decrease activity than someone giving the same amount of kudos to people with higher activity

## Evaluation of *behavior*

-   behavior = *Z*

-   network = *X*

-   X ij is outdegrees

-   1 if there is relation

-   0 if there is no relation

-   we want alter mean effect which is the alter TOTAL effect DIVIDED by MEAN effect

-   RSiena models the scores before doing the formulas so the RSiena manual likely differs from what we talked about

# Netw. Statistics: Reciprocity

## RSiena statistics

SBeh = behavior

SNet = network

## Reciprocity formula form RSiena manual

<https://www.stats.ox.ac.uk/~snijders/siena/RSiena_Manual.pdf>

Xij = value for tie from i to j [1 or 0]

Xji = value tie from j to i [1 or 0]

# Final project information

-   statistics: include degrees, reciprocity

-   control for skewness (indegree, outdegree, popularity)

often: friends of friends become friends. Most often bc of *opportunity structures*.

-   including transitivity effect

    -   only include one transtrip effect transtrip1 not the other ones for now

# GOF

- tests if our model is able to replicate the data

# Playing arond with my data

##clean workspace

```{r}
# start with clean workspace
rm(list = ls())
```

## packages

```{r}
#install.packages("tidyverse")
require (tidyverse)
require (RSiena)
```

## Functions

```{r}

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
```

## KEEP THIS ON TOP OF YOUR RSIENA CODE ALWAYS!

```{r}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids, ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```

## Load data

```{r}
scholars <- fload("./data/processed/scholars_20240924.rda")
```

```{r}
#save the output of your function
test <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

```

### demographics    
```{r}

df_ego_demo <- test$data$demographics
df_ego_demo <- do.call(rbind.data.frame, df_ego_demo)

df_ego_oa <- test$data$scholars_oa 
df_ego_oa <- do.call(rbind.data.frame, df_ego_oa)
# 
# df_ego_works <- test$data$works  #did not do anything with this one yet
# 
df_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))

```

```{r}
df_ego_demo <- do.call(rbind.data.frame, df_ego_demo)
df_ego_demo[[1]]

str(df_ego_oa, 1)
# df_ego_oa <- do.call(rbind.data.frame, df_ego_oa)
# df_ego_oa[[1]]

# ?join.data.frame

# rownames(nets[1,,]) == df_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
# Yes, the correct ones!!
```

```{r}
df_ego$Naam
df_ego$gender <- c(0,1,0,1,0,0,1,1,0,0,
                   0,1,1,0,1,0,1,1,1,1,
                   0,1,1,0,1,1,0,1,0,1,
                   0,1,0,0,0,0,0,0,1,0,
                   0,1,1,1,0,1,0,1,0,0) 
# manually filled in gender 0 = woman, 1 = man, based on majority NL voornamen bank

# view(df_ego[c('Naam', 'gender', 'au_id')])

```

# RSiena

1.  Define data
2.  Create effects structure
3.  Get initial description
4.  Specify model
5.  Estimate model
6.  GOF measures [week7]

## Step 1: Define data

### dep var

```{r}
nets <- test$nets
wave1 <- test$nets[1, , ]
wave2 <- test$nets[2, , ]

# let us put the diagonal to zero
diag(wave1) <- 0
diag(wave2) <- 0
# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
# put the nets in an array
net_soc_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

# dependent
net <- sienaDependent(net_soc_array)

```

### gender

```{r}

gender <- coCovar(df_ego$gender)

```

### combine data

```{r}
mydata <- sienaDataCreate(net, gender)
```

## Step 2: Create effects structure

```{r}
myeff <- getEffects(mydata)
```

## Step 3: Get initial description

## Step 4: Specify model

```{r}
myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, sameX, interaction1 = 'gender')
```

## Step 5: Estimate model

```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
ansM1 <- siena07(myAlgorithm, data = mydata, effects = myeff, returnDeps = TRUE)

#0 and 1 is about whether tie is the same now, notabout male and female anymore

```

## (Step 6: GOF measures [week7]