### 2024-10-30 clarifyRQsandDesign.Rmd
restructurecommit

---
title: "Clarify RQ and Design"
author: "Verooo"
date: "2024-10-30"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Clarify Research Questions

**RQ1a:**

To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Political Science Department at Radboud University?

aka. looking at whole sample: do we see homophiliy? how much ?

**RQ1b:**

To what extent do the two departments differ in their gender homophily?

aka. looking at Sociology Department and extent of homphily in there \
and comparing it with Political Science Department and extent of homophily in there.

**RQ2:**

Jochem's input: insert here another RQ before last RQ

Do we observe gender homophily (preferences) in collaboration networks...

Questions:

-   Does this question need to ask whether we observe gender homophily in collaboration networks to begin with?

    -   Isn't that the same question as RQ1a then?

-   Does the question need to ask whether we observe gender homophily in collaborations over time?

    -   Did not intent to look at changes over time. this would change things up.

-   Or should it be about the prestige variable in some way? \
    e.g. whether we observe gender homophily within collaboration networks of high prestige scholars and within collaboration networks of lower prestige scholars

    -   but that presupposes that we have pre-established that there are collaboration networks of high-prestige scholars and low prestige scholars

    -   also doesn't this question take away from RQ3?

-   Concern: Overall, I want to keep things simple. Adding a RQ might be needed to help set up the last RQ here but I wonder if there is a way to simplify the project design then?

**RQ3:**

How to does gender homophily affect collaboration differently for collaborations between staff who have published in low vs. high prestige journals?

aka. looking at whole sample: does the gender homophily in collaboration networks look different depending on the scholars' publishing prestige?

# Design

# Project parts

## Research Design

-   Clarify Research Questions

-   Clarify design to answer this question

-   Clarify which data is needed in the design

## Steps to analysis

-   Step 1: Define data

-   Step 2: Create effects structure

-   Step 3: Get initial description

-   Step 4: Specify model

-   Step 5: Estimate model

-   Step 6: Goodness of Fit

## Report writing

-   Introduction

-   Theory

-   Data & Methods

-   Results

-   Discussion & Conclusion