### 2024-11-03 1definedata.Rmd
defindedatapolsci

---
title: "1-Define Data"
author: "Verooo"
date: "2024-11-01"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
klippy::klippy(color = 'darkred')
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

[**Defining Data**]{.underline}

# Before the start

## Clean workspace

```{r}
# start with clean workspace
rm(list = ls())

```

## Packages

```{r}
# install.packages('tidyverse')
require(tidyverse)

# install an old version of RSiena including sienaRI
#remove.packages("RSiena") #use this to remove the current version of RSiena
#devtools::install_version("RSiena", version = "1.3.14", repos = "C:/Users/Vero/OneDrive - Radboud Universiteit/Documents/GitHub/labjournal") # This installs an older version

# Run R as administrator when you install the older version!!

# You should end up with version 1.3.14 loaded in instead of 1.4.7, 
# but then make sure there isn't a install.packages("RSiena") function running somewhere, because then it will reinstall the newest version

# require(RSiena)
# require(RsienaTwoStep)

```

I do not seem to be able to call on the package RSiena, nor can I reinstall it.... I know programming never works on the first try but... goddamnit, why not even after 15?

## Functions

### Question

**HOW do I get the discipline of political science in there too?**

**How can I find out what the name for the discipline of political science in the original data frame (e.g. whether it is called polsci, political science, or politicalscience)?**

**I don't know how to view the data set BEFORE we include only those scholars that have discipline = sociology. I cannot view the data**

I tried removing the discipline = "sociology" part in the f_pubnets, fcolnet and all following functions that select data BUT SINCE 'discipline' is also relevant in the rest of the functions' code, it does not end up doing what I hoped (having a dataset with scholars of no specific discipline because then I could hopefully check how the department is named in the data and then find out how to add that to the data set initially).

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

## more functions

```{r}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```

## load and inspect data

```{r}
scholars <- fload("./data/processed/scholars_20240924.rda") #make sure to lead fload first ofcourse

#inspecting scholars_20240924.rda object
#df_ego <- do.call(rbind.data.frame, scholars$demographics)
#unique(df_ego$discipline.22)

# save the output of your function
testsoc  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)),

                            type = c("first"))

# I DONT KNOW HOW TO MAKE ALL SCHOLARS BE PART OF SAME FUNCTION ("sociology"+ "political science" does not work...)

testpolsci <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "political science", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

------------------------------------------------------------------------

# Define Data

## waves - sociology department

```{r}

# Jochem messed up and did: waves, rows, columns [1, , ]
socwave1 <- testsoc$nets[ , ,1]  #rows, columns, waves
socwave2 <- testsoc$nets[ , ,2]  #rows, columns, waves
# let us put the diagonal to zero
diag(socwave1) <- 0
diag(socwave2) <- 0
# we want a binary tie (not a weighted tie)
socwave1[socwave1 > 1] <- 1
socwave2[socwave2 > 1] <- 1
# put the nets in an array
net_soc_array <- array(data = c(socwave1, socwave2), dim = c(dim(socwave1), 2))

str(net_soc_array)  #now we basically restructured this 
# output: num [1:50, 1:50, 1:2] 0 0 0 0 0 0 0 0 0 0
```

## waves - polsci department

```{r}
# Jochem messed up and did: waves, rows, columns [1, , ]
polwave1 <- testpolsci$nets[ , ,1]  #rows, columns, waves
polwave2 <- testpolsci$nets[ , ,2]  #rows, columns, waves
# let us put the diagonal to zero
diag(polwave1) <- 0
diag(polwave2) <- 0
# we want a binary tie (not a weighted tie)
polwave1[polwave1 > 1] <- 1
polwave2[polwave2 > 1] <- 1
# put the nets in an array
net_pol_array <- array(data = c(polwave1, polwave2), dim = c(dim(polwave1), 2))

str(net_pol_array)  #now we basically restructured this

```

### **QUESTION:**

**What's the difference between putting the 2 and the 1 at the dimensions. I feel like one is appropriate since I am only looking at one wave, not two, but since the output is among the 'values' not 'data', I cannot view the variable net_soc_array(1) nor "visualize" what that means in terms of its structure. I don't really know what the difference would be between the 1:2 and the 1 in the outputs, structurally.**

### ANSWER:

RSiena needs at least two waves, because it models the evolution of the networks. We have to store the networks in an array before we can convert it to an dependent variable objects. Moreover, for RSiena the dimensions should be in the following order: rows, columns, waves.

I 'messed up' my function (for didactic reasons ;-) )and stored the nets in the following order: waves, rows, columns. Thus, either correct my function or retrieve the separate waves and store them in the order RSiena requires.

### QUESTION TO THE ANSWER:

We had agreed on me only investigating one time point for the sake of simplicity. It seems I basically have no choice but to investigate two waves, revisit my RQs and make an additional RQ about tie changes over time, right?

Then I would end up with 4 RQs and that seems like more than it needs to be for this project (great for a research incentive, but for this class I need to keep it simple, as I am currently very overworked).

Any advice on how I can keep it simple, nonetheless?

## Examine structure

```{r}
dim(testsoc$nets)
```

```{r}
dim(testpolsci$nets)
```

## My data

### sociology dep.

not using net_soc_array for now since I am unsure about its use due to question above

```{r}
# dependent
# net <- sienaDependent(net_soc_array)

# dependent 
nets <- testsoc$nets

df_ego_demo <- testsoc$data$demographics
df_ego_oa <- testsoc$data$scholars_oa
df_ego_works <- testsoc$data$works

# DEMO
  str(df_ego_demo, 1)

  df_ego_demo <- do.call(rbind.data.frame, df_ego_demo)
  df_ego_demo[[1]]

# OA
  #str(df_ego_oa, 1)
  df_ego_oa[[1]]
  df_ego_oa <- do.call(rbind.data.frame, df_ego_oa)

  # --> EGO DATAFRAME with demo and oa (sociology)
  ?join.data.frame
  df_soc_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))

  rownames(nets[1,,]) == df_soc_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
  # --> yes, the correct ones!!!

  df_ego_works[[1]]$author 
#authors of all the works that first scholar (here: Ana Macanovic) is part of

```

### polsci department

```{r}
# dependent
# pol_net <- sienaDependent(net_pol_array)

# dependent 
pol_nets <- testpolsci$nets

df_pol_ego_demo <- testpolsci$data$demographics
df_pol_ego_oa <- testpolsci$data$scholars_oa
df_pol_ego_works <- testpolsci$data$works

# DEMO
  str(df_pol_ego_demo, 1)

  df_pol_ego_demo <- do.call(rbind.data.frame, df_pol_ego_demo)
  df_pol_ego_demo[[1]]

# OA
  #str(df_pol_ego_oa, 1)
  df_pol_ego_oa[[1]]
  df_pol_ego_oa <- do.call(rbind.data.frame, df_pol_ego_oa)

# --> EGO DATAFRAME with demo and oa (political science)
  df_pol_ego <- left_join(df_pol_ego_demo, df_pol_ego_oa, by = join_by(au_id == id))

  rownames(pol_nets[1,,]) == df_pol_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
  # --> yes, the correct ones!!!
```

## Gender

### gender-soc

```{r}
df_soc_ego$Naam # list of all names in sociology department

df_soc_ego$gender <- c(0,1,0,1,0,0,1,1,0,0,
                       0,1,1,0,1,0,1,1,1,1,
                       0,1,1,0,1,1,0,1,0,1,
                       0,1,0,0,0,0,0,0,1,0,
                       0,1,1,1,0,1,0,1,0,0) 
# manually filled in gender 0 = woman, 1 = man, based on majority given in NL voornamen bank

# check if it went smoothly 
view(df_soc_ego[c('Naam', 'gender', 'au_id')])

```

### gender - polsci

```{r}
df_pol_ego$Naam # list of all names

df_pol_ego$gender <- c(1,0,1,0,0,0,1,0,1,0,
                       1,1,0,0,1,1,1,0,0,1,
                       0,1,1,0,0,1,1,1,0,0,
                       0,0,0,1,1,0,0,1,1,0,
                       0,0,0,1,0,1,1,1,0,0,0) 
    # manually filled in gender 0 = woman, 1 = man, based on majority given in NL voornamen bank, Guangyu was not in name bank but based on the employee-picture of RU, this person presents feminine, thus I assumed 0.

# check if it went smoothly 
view(df_pol_ego[c('Naam', 'gender', 'au_id')])

#covariate 
# gender <- coCovar(df_ego$gender)
```

### joining polsci and soc data

```{r}
df_both_ego <- full_join(df_soc_ego, df_pol_ego, by = join_by
                    (au_id == au_id, 
                     Naam==Naam, 
                     gender == gender, 
                     discipline.22 == discipline.22,
                     discipline.24 == discipline.24,
                     counts_by_year== counts_by_year, 
                     works_count == works_count, 
                     cited_by_count == cited_by_count, 
                     Universiteit1.22 == Universiteit1.22,
                     Universiteit2.22 == Universiteit2.22,
                     Universiteit1.24 == Universiteit1.24,
                     Universiteit2.24 == Universiteit2.24,
                     Universiteit.22 == Universiteit.22, 
                     Universiteit.24 == Universiteit.24,
                     no_oa_id == no_oa_id,
                     display_name == display_name,
                     display_name_alternatives == display_name_alternatives,
                     ids== ids,
                     orcid == orcid,
                     year.22 == year.22,
                     year.24 == year.24,
                     Functie.24== Functie.24,
                     relevance_score == relevance_score,
                     affiliation_display_name == affiliation_display_name,
                     affiliation_ror == affiliation_ror,
                     affiliation_id == affiliation_id,
                     affiliation_country_code == affiliation_country_code,
                     affiliation_type == affiliation_type,
                     affiliations_other == affiliations_other,
                     affiliation_lineage == affiliation_lineage,
                     Specialisatie.22 == Specialisatie.22,
                     Specialisatie.24 == Specialisatie.24,
                     x_concepts == x_concepts,
                     works_api_url == works_api_url))
# 101 scholars (50 sociology + 51 polsci) and 43 variables (as opposed to 21 bc i could not join them all.)

# GODDAMMIT WHY DO I ONLY FIND THIS NOW?!?!?!?!
# mydata <- sienaDataCreate(net, gender)
```

## QUESTION:

With 101 scholars, instead of the nr. of scholars of looking at just one department, I wonder if this is where we could simplify. Looking at a change over time, but not looking through a lens that wants to compare the two departments. Instead, I could just look at one department.

What do you think?

## H-index webscraping

Credits to Lucan for the code.

```{r}
# Add h-index and i10-index columns to df_ego
df_ego <- df_both_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_)

# Get H-index and i10-index 
get_openalex_metrics <- function(openalex_id) {
  url <- paste0("https://api.openalex.org/authors/", openalex_id)
  author_data <- tryCatch({
    jsonlite::fromJSON(url)
  }, error = function(e) {
    message(paste("Error fetching data for OpenAlex ID:", openalex_id))
    return(NULL)
  })

  if (!is.null(author_data)) {
    h_index <- author_data$summary_stats$h_index
    i10_index <- author_data$summary_stats$i10_index
    works_count <- author_data$works_count
    cited_by_count <- author_data$cited_by_count
    return(list(h_index = h_index, i10_index = i10_index, works_count = works_count, cited_by_count = cited_by_count))
  } else {
    return(list(h_index = NA, i10_index = NA, works_count = NA, cited_by_count = NA))
  }
}

# Loop for each scholar 
for (i in 1:nrow(df_ego)) {
  openalex_id <- df_ego$au_id[i]  # assuming au_id is the OpenAlex ID in df_ego
  metrics <- get_openalex_metrics(openalex_id)
  df_ego$h_index[i] <- metrics$h_index
  df_ego$i10_index[i] <- metrics$i10_index
  df_ego$works_count[i] <- metrics$works_count
  df_ego$cited_by_count[i] <- metrics$cited_by_count
}

# View updated df_ego with h-index and i10-index
view(df_ego[c('au_id','Naam','gender','h_index','i10_index')])

#covariate 
# h_index <- coCovar(df_ego$h_index)
```

## Actually define data:

```{r}
# dependent:
# net <- sienaDependent(net_soc_array)

#covariate 1
# gender <- coCovar(df_ego$gender)
# covariate 2
# h_index <- coCovar(df_ego$h_index)

#combining data: wish I knew of this shit earlier...
# mydata <- sienaDataCreate(net, gender)
```

# Create effects structure

```{r}
#myeff <- getEffects(mydata)
```