### 2024-11-07 revisitwithJochem.Rmd
checkinwithjochempush

---
title: "Plotting plots and Hotzenplotz"
author: "Verooo"
date: "2024-11-05"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Meeting with Jochem

look at publications website (sunbelt workshop)

# Getting started

```{r clean}
rm(list = ls())
```

## Packages

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)

require(ggraph)
#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

## More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

## Load data

```{r data}
scholars <- fload("./data/processed/scholars_20240924.rda") #make sure to lead fload first ofcourse
```

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("all")) # all, first, last # which ties in there

# USE not all but first or last so the ties are directed from "perspective" of first author of the publications 
# usually first most prestige

```

# Gender

## saving data for each deapartment

```{r}
#inspecting scholars_20240924.rda object
#df_ego <- do.call(rbind.data.frame, scholars$demographics)
#unique(df_ego$discipline.22)

# save the output of your function
testsoc  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)),

                            type = c("first"))

# I DONT KNOW HOW TO MAKE ALL SCHOLARS BE PART OF SAME FUNCTION ("sociology"+ "political science" does not work...)

testpolsci <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "political science", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

#### soc department

```{r}
# # dependent
# net <- sienaDependent(net_soc_array)

# dependent ?
# nets <- testsoc$nets

df_ego_demo <- testsoc$data$demographics
df_ego_oa <- testsoc$data$scholars_oa
df_ego_works <- testsoc$data$works

# DEMO
  str(df_ego_demo, 1)

  df_ego_demo <- do.call(rbind.data.frame, df_ego_demo)
  df_ego_demo[[1]]

# OA
  #str(df_ego_oa, 1)
  df_ego_oa[[1]]
  df_ego_oa <- do.call(rbind.data.frame, df_ego_oa)

  # --> EGO DATAFRAME with demo and oa (sociology)
  ?join.data.frame
  df_soc_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))

  rownames(nets[1,,]) == df_soc_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
  # --> yes, the correct ones!!!

  df_ego_works[[1]]$author 
#authors of all the works that first scholar (here: Ana Macanovic) is part of

```

#### polsci department

```{r}
# dependent
# pol_net <- sienaDependent(net_pol_array)

# dependent 
pol_nets <- testpolsci$nets

df_pol_ego_demo <- testpolsci$data$demographics
df_pol_ego_oa <- testpolsci$data$scholars_oa
df_pol_ego_works <- testpolsci$data$works

# DEMO
  str(df_pol_ego_demo, 1)

  df_pol_ego_demo <- do.call(rbind.data.frame, df_pol_ego_demo)
  df_pol_ego_demo[[1]]

# OA
  #str(df_pol_ego_oa, 1)
  df_pol_ego_oa[[1]]
  df_pol_ego_oa <- do.call(rbind.data.frame, df_pol_ego_oa)

# --> EGO DATAFRAME with demo and oa (political science)
  df_pol_ego <- left_join(df_pol_ego_demo, df_pol_ego_oa, by = join_by(au_id == id))

  rownames(pol_nets[1,,]) == df_pol_ego$au_id # checking if the author id's are the same as in df_ego in nets dataframe 
  # --> yes, the correct ones!!!
```

### gender-soc

```{r gender s}
df_soc_ego$Naam # list of all names in sociology department

df_soc_ego$gender <- c(0,1,0,1,0,0,1,1,0,0,
                       0,1,1,0,1,0,1,1,1,1,
                       0,1,1,0,1,1,0,1,0,1,
                       0,1,0,0,0,0,0,0,1,0,
                       0,1,1,1,0,1,0,1,0,0) 
# manually filled in gender 0 = woman, 1 = man, based on majority given in NL voornamen bank

# check if it went smoothly 
view(df_soc_ego[c('Naam', 'gender', 'au_id')])

# fsave(df_soc_ego)

```

### gender - polsci

```{r gender p}
df_pol_ego$Naam # list of all names

df_pol_ego$gender <- c(1,0,1,0,0,0,1,0,1,0,
                       1,1,0,0,1,1,1,0,0,1,
                       0,1,1,0,0,1,1,1,0,0,
                       0,0,0,1,1,0,0,1,1,0,
                       0,0,0,1,0,1,1,1,0,0,0) 
    # manually filled in gender 0 = woman, 1 = man, based on majority given in NL voornamen bank, Guangyu was not in name bank but based on the employee-picture of RU, this person presents feminine, thus I assumed 0.

# check if it went smoothly 
view(df_pol_ego[c('Naam', 'gender', 'au_id')])

#covariate 
# gender <- coCovar(df_ego$gender)

# fsave(df_pol_ego)
```

### save gender data sets

```{r gender data}
soc_gender <- fload("./data/processed/df_soc_ego_20241105.rda")
pol_gender <- fload("./data/processed/df_pol_ego_20241105.rda")
```

### Combine data (gender)

```{r combine gender set}
df <- test$data
df_ego <- do.call(rbind.data.frame, df$demographics)

soc_gender <- soc_gender[,c("Naam", "gender")]
pol_gender <- pol_gender[,c("Naam", "gender")]

gender <- rbind(soc_gender, pol_gender) # combining data sets

df_ego <- df_ego %>% 
            left_join(gender) #dataset 
```

# Quick RSiena

```{r RSiena}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

#put diagonals to zero 
diag(wave1) <- 0
diag(wave2) <- 0
# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
# put the nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

gender <- df_ego$gender

#dependent
nets <- sienaDependent(net_array)
#covariate
genderrs <- coCovar(gender)
#define data 
mydata <- sienaDataCreate(nets, genderrs)

#define effects structure 
myeff <- getEffects(mydata)

#skipped description

myeff <- includeEffects(myeff, isolateNet, inPop, outAct)

myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1 = "genderrs")

#look up which effects to iclude

myAlgorithm <- sienaAlgorithmCreate(modelType = c(nets=6), projname = "proProject")
# undirected: model 6 "additive model", #check manual for this

# estimate
model1 <- siena07(myAlgorithm, data = mydata, effects = myeff, returnDeps = TRUE)
#latter needed for GOF

model1

# output:  convergence ration NA probably bc we chose all ties to be included (type = "all") earlier
# we can go back and change that or change it here in the myAlgorithm function

# Make it a directed network 
# focus: writing, h index, 
# model interpretation

```

# H-index webscraping

Credits to Lucan for the code.

```{r h-index, include=FALSE, eval= FALSE}
# Add h-index and i10-index columns to df_ego
df_ego <- df_both_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_)

# Get H-index and i10-index 
get_openalex_metrics <- function(openalex_id) {
  url <- paste0("https://api.openalex.org/authors/", openalex_id)
  author_data <- tryCatch({
    jsonlite::fromJSON(url)
  }, error = function(e) {
    message(paste("Error fetching data for OpenAlex ID:", openalex_id))
    return(NULL)
  })

  if (!is.null(author_data)) {
    h_index <- author_data$summary_stats$h_index
    i10_index <- author_data$summary_stats$i10_index
    works_count <- author_data$works_count
    cited_by_count <- author_data$cited_by_count
    return(list(h_index = h_index, i10_index = i10_index, works_count = works_count, cited_by_count = cited_by_count))
  } else {
    return(list(h_index = NA, i10_index = NA, works_count = NA, cited_by_count = NA))
  }
}

# Loop for each scholar 
for (i in 1:nrow(df_ego)) {
  openalex_id <- df_ego$au_id[i]  # assuming au_id is the OpenAlex ID in df_ego
  metrics <- get_openalex_metrics(openalex_id)
  df_ego$h_index[i] <- metrics$h_index
  df_ego$i10_index[i] <- metrics$i10_index
  df_ego$works_count[i] <- metrics$works_count
  df_ego$cited_by_count[i] <- metrics$cited_by_count
}

# View updated df_ego with h-index and i10-index
view(df_ego[c('au_id','Naam','gender','h_index','i10_index')])

#covariate 
# h_index <- coCovar(df_ego$h_index)
```

# Plotting

```{r}
test_w1 <- igraph::graph_from_adjacency_matrix(
  test$nets[1,,], # first wave of data. 
  # (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL)

plot(test_w1,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)
```

```{r}
names(df_ego) #see labels of variables

plot(test_w1,
vertex.color = ifelse(df_ego$gender == 1, "red", "blue"), #now, I can use actor attributes for plotting.
vertex.label = NA,
edge.width = 0.2,
edge.arrow.size =0.2)

# if you dont want plot to turn, save the layout (look at SNASS)
```