### 2024-11-09 weekbacktrack.Rmd

---
title: "week getbackontrack"
author: "Verooo"
date: "2024-10-27"
output: html_document
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Brainstorm - Getting back on track

## Project parts

### Research Design

-   Clarify Research Questions

-   Clarify design to answer this question

-   Clarify which data is needed in the design

### Steps to analysis

-   Step 1: Define data

-   Step 2: Create effects structure

-   Step 3: Get initial description

-   Step 4: Specify model

-   Step 5: Estimate model

-   Step 6: Goodness of Fit

### Report writing

-   Introduction

-   Theory

-   Data & Methods

-   Results

-   Discussion & Conclusion

### Simultaneously possible

-   Clarify RQs and Design \<-\> Writing introduction and theory

-   Defining Data & effects structure, initial description and model specification \<-\> Data & Methods part

-   Estimate Model & Goodness of fit \<-\> Results section

-   Afterwards: Discussion and Conclusion

## Timeline

tentative deadline: Saturday, 23.11.24

+--------------------------------------------------------------------+------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------+------------------------------------------+
| Mon                                                                | Tue                                            | Wed                                                           | Thu                                                                    | Fri                                                                      | Sat                                                                                     | Sun                                      |
+:===================================================================+:===============================================+:==============================================================+:=======================================================================+:=========================================================================+:========================================================================================+:=========================================+
| 28\.                                                               | 29\.                                           | 30\.                                                          | 31\.                                                                   | 01.11.                                                                   | 02\.                                                                                    | 03\.                                     |
|                                                                    |                                                |                                                               |                                                                        |                                                                          |                                                                                         |                                          |
| ![](images/used_arson%20frog.jpg)                                  | ![](images/used_arson%20frog.jpg){width="952"} | ![](images/socialnetworking.jpg){width="800"}Clarify RQs [D]\ | ![](images/socialnetworking_meh.jpg){width="800"}Intro and theory [W]\ | ![](images/socialnetworking.jpg){width="800"}Intro and theory [W]\       | ![](images/socialnetworking_meh.jpg){width="800"}Define data [S1]                       | ![](images/Trixie-joke.jpg){width="800"} |
|                                                                    |                                                | & Clarify design [D]                                          |                                                                        | Clarify [D] & define data [S1]                                           |                                                                                         |                                          |
+--------------------------------------------------------------------+------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------+------------------------------------------+
| 04\.                                                               | 05\.                                           | 06\.                                                          | 07\.                                                                   | 08\.                                                                     | 09\.                                                                                    | 10\.                                     |
|                                                                    |                                                |                                                               |                                                                        |                                                                          |                                                                                         |                                          |
| ![](images/socialnetworking_meh.jpg){width="800"}Define data [S1]\ | ![](images/used_arson%20frog.jpg){width="952"} | ![](images/used_arson%20frog.jpg)                             | ![](images/socialnetworking_meh.jpg){width="800"}\                     | ![](images/socialnetworking.jpg){width="800"}Initial description [S3]\   | ![](images/socialnetworking.jpg){width="800"}Plotting                                   | ![](images/Trixie-joke.jpg){width="800"} |
| & create effects structure [S2]                                    |                                                |                                                               | Create effects structure [S2]\                                         | & plotting                                                               |                                                                                         |                                          |
+--------------------------------------------------------------------+------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------+------------------------------------------+
| 11\.                                                               | 12\.                                           | 13\.                                                          | 14\.                                                                   | 15\.                                                                     | 16\.                                                                                    | 17\.                                     |
|                                                                    |                                                |                                                               |                                                                        |                                                                          |                                                                                         |                                          |
| ![](images/socialnetworking_meh.jpg){width="800"}Plotting          | ![](images/used_arson%20frog.jpg){width="952"} | ![](images/used_arson%20frog.jpg)                             | ![](images/socialnetworking.jpg){width="800"}Specify model [S4]\       | ![](images/socialnetworking.jpg){width="800"}Data & methods [W]          | ![](images/socialnetworking.jpg){width="800"}Estimate model [S5] & goodness of fit [S6] | ![](images/Trixie-joke.jpg){width="800"} |
|                                                                    |                                                |                                                               | Data & methods part [W]                                                |                                                                          |                                                                                         |                                          |
+--------------------------------------------------------------------+------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------+------------------------------------------+
| 18\.                                                               | 19\.                                           | 20\.                                                          | 21\.                                                                   | 22\.                                                                     | **23.**                                                                                 | 24\.                                     |
|                                                                    |                                                |                                                               |                                                                        |                                                                          |                                                                                         |                                          |
| ![](images/socialnetworking_meh.jpg){width="800"}Results [W]       | ![](images/used_arson%20frog.jpg){width="952"} | ![](images/used_arson%20frog.jpg)                             | ![](images/socialnetworking_meh.jpg){width="800"}Finish results [W]    | ![](images/socialnetworking.jpg){width="800"}Discussion & conclusion [W] | ![](images/socialnetworking.jpg){width="800"}Read over project & hand in!               | ![](images/Trixie-joke.jpg){width="800"} |
+--------------------------------------------------------------------+------------------------------------------------+---------------------------------------------------------------+------------------------------------------------------------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------+------------------------------------------+

: Days overview

**Legend**

::: columns
::: {.column width="50%"}
![social network plot (color) = time for SNA (obvious... or not?)](images/socialnetworking.jpg){width="80"}

![social network plot (muted) = not a lot of time for SNA, but SOME](images/socialnetworking_meh.jpg){width="80"}
:::

::: {.column width="50%"}
![arson = otherwise busy/NO time for SNA](images/used_arson%20frog.jpg){width="80"}

![you think this is a joke? = chores, admin work, and mental breakdowns/ NO time for SNA](images/Trixie-joke.jpg){width="80"}
:::
:::

[D] = task refers to research design

[S#] = task refers to step of analysis

[W] = task refers to writing the report

### Timeline Question

When is plotting done in this progression? At the initial description part or earlier?

## Project Planning

I have decided to drop another course I was taking and focus on the deadline of this course.

I think the workload of both courses' deadlines exceeds my capacities, especially since the courses of PER 3 already start on 04.11. and I also have duties relating to my board year simultaneously.

### Suggestions for Jochem:

I will update you once a week to inform you of my progress. This can act as an "accountability-measure" and pushes me to face some struggles head-on, that I would otherwise try to avoid. It also creates a predesignated time to discuss the project, which can help me ask questions I would otherwise deem "not important enough to reach out to you".\
I propose that I write you an email about the **current state of affairs on Mondays**, since I cannot work on the project on Tuesday and Wednesday anyways and that gives you some time to reply. Please let me know if this is okay for you.\
In your reply I would appreciate input on how to move on in case I am facing a hurdle that "stalls" the project's progress, and I also welcome redirection if I seem to get "lost in the sauce". Any hints on how to take "shortcuts" that I was not aware of are also very much invited (e.g. "I know of this online resource that is about *exactly* this issue.").

Thank you in advance.

## ACTUAL Timeline

deadline: Saturday, 23.11.24

+-----------------------------------------------------------------------------------------------+------------------------------------------------------------+---------------------------------------------------------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------------------+---------------------------------------------------------------------+
| Mon                                                                                           | Tue                                                        | Wed                                                           | Thu                                                                                        | Fri                                                             | Sat                                                                       | Sun                                                                 |
+:==============================================================================================+:===========================================================+:==============================================================+:===========================================================================================+:================================================================+:==========================================================================+:====================================================================+
| 28\.                                                                                          | 29\.                                                       | 30\.                                                          | 31\.                                                                                       | 01.11.                                                          | 02\.                                                                      | 03\.                                                                |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| ![](images/used_arson%20frog.jpg)                                                             | ![](images/used_arson%20frog.jpg){width="952"}             | ![](images/socialnetworking.jpg){width="800"}Clarify RQs [D]\ | ![](images/socialnetworking_meh.jpg){width="800"}Rework schedule [according Jochems input] | ![](images/socialnetworking.jpg){width="800"}\                  | ![](images/timetopanic.gif){width="800"}[chores at home, depression day]  | ![](images/socialnetworking_meh.jpg)Get polSci data,                |
|                                                                                               |                                                            | [Asked Jochem for input]                                      |                                                                                            | Clarify [D]\                                                    |                                                                           |                                                                     |
|                                                                                               |                                                            |                                                               | OUTLINE Intro and theory [W]\                                                              | & define data [S1]\                                             |                                                                           | define data                                                         |
|                                                                                               |                                                            |                                                               | (missing H3 still)                                                                         |                                                                 |                                                                           |                                                                     |
|                                                                                               |                                                            |                                                               |                                                                                            | \* struggle with polsci data \*                                 |                                                                           |                                                                     |
+-----------------------------------------------------------------------------------------------+------------------------------------------------------------+---------------------------------------------------------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------------------+---------------------------------------------------------------------+
| 04\.                                                                                          | 05\.                                                       | 06\.                                                          | 07\.                                                                                       | 08\.                                                            | 09\.                                                                      | 10\.                                                                |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| ![](images/socialnetworking_meh.jpg){width="800"}\                                            | ![](images/used_arson%20frog.jpg){width="952"}[work]       | ![](images/used_arson%20frog.jpg)[university]                 | ![](images/socialnetworking_meh.jpg){width="800"}\                                         | ![](images/socialnetworking.jpg){width="800"}\                  | ![](images/socialnetworking.jpg){width="800"}rewrite intro and Theory [W] | ![](images/timetopanic.gif){width="800"}[Grandmas bday celebration] |
| fix package error,                                                                            |                                                            |                                                               | Read Jochems answer & revisit code                                                         | fix bugging webjournal                                          |                                                                           |                                                                     |
|                                                                                               | Talk to Jochem                                             |                                                               |                                                                                            |                                                                 | read into specifying model [S4]                                           |                                                                     |
| Define variables (get stuck with sienaDependent)                                              |                                                            |                                                               | & create effects structure [S2]                                                            | journal, make submenus                                          |                                                                           |                                                                     |
|                                                                                               | [university]                                               |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| check in with Jochem                                                                          |                                                            |                                                               | make h index                                                                               | Reorganize code                                                 |                                                                           |                                                                     |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| Plotting                                                                                      |                                                            |                                                               |                                                                                            | Starting with initial description [S3]                          |                                                                           |                                                                     |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
|                                                                                               |                                                            |                                                               |                                                                                            | Plotting                                                        |                                                                           |                                                                     |
+-----------------------------------------------------------------------------------------------+------------------------------------------------------------+---------------------------------------------------------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------------------+---------------------------------------------------------------------+
| 11\.                                                                                          | 12\.                                                       | 13\.                                                          | 14\.                                                                                       | 15\.                                                            | 16\.                                                                      | 17\.                                                                |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| ![](images/socialnetworking_meh.jpg){width="800"}Write Data & methods part [W]                | ![](images/used_arson%20frog.jpg){width="952"}[university] | ![](images/used_arson%20frog.jpg)[university]                 | ![](images/socialnetworking.jpg){width="800"}Data & methods part [W]\                      | ![](images/socialnetworking.jpg){width="800"}estimate model     | ![](images/socialnetworking.jpg){width="800"}goodness of fit [S6]         | ![](images/kermit-panic.gif)                                        |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| Check in with Jochem                                                                          |                                                            |                                                               |                                                                                            | goodness of fit                                                 | Start Results [W]                                                         | [chores at home and university]                                     |
+-----------------------------------------------------------------------------------------------+------------------------------------------------------------+---------------------------------------------------------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------------------+---------------------------------------------------------------------+
| 18\.                                                                                          | 19\.                                                       | 20\.                                                          | 21\.                                                                                       | 22\.                                                            | **23.**                                                                   | 24\.                                                                |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| ![](images/used_arson%20frog.jpg)[Research project course starts: little time to work on SNA] | ![](images/used_arson%20frog.jpg){width="952"}[university] | ![](images/used_arson%20frog.jpg)[university]                 | ![](images/socialnetworking_meh.jpg){width="800"}Results [W]                               | ![](images/socialnetworking.jpg){width="800"}Finish Results [W] | ![](images/socialnetworking.jpg){width="800"}Discussion & conclusion [W]  |                                                                     |
|                                                                                               |                                                            |                                                               |                                                                                            |                                                                 |                                                                           |                                                                     |
| Check in with Jochem?                                                                         |                                                            |                                                               |                                                                                            | Discussion & conclusion [W]                                     | ![](images/kermit-panic.gif){width="500"}Format project website& hand in! |                                                                     |
+-----------------------------------------------------------------------------------------------+------------------------------------------------------------+---------------------------------------------------------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------------------+---------------------------------------------------------------------+

: Days overview

**Legend**

::: columns
::: {.column width="50%"}
![social network plot (color) = time for SNA (obvious... or not?)](images/socialnetworking.jpg){width="80"}

![social network plot (muted) = not a lot of time for SNA, but SOME](images/socialnetworking_meh.jpg){width="80"}

![arson = otherwise busy/NO time for SNA](images/used_arson%20frog.jpg){width="80"}
:::

::: {.column width="50%"}
![time to panic = chores, admin work, and mental breakdowns/ NO time for SNA](images/timetopanic.gif){width="80"}

![kermit panicking = self-explanatory](images/kermit-panic.gif){width="80"}
:::
:::

[D] = task refers to research design

[S#] = task refers to step of analysis

[W] = task refers to writing the report