### 2024-11-16 draftintro.Rmd
darkmodeleggo

---
title: "Daft-Introduction"
author: "Verooo"
date: "2024-10-31"
output: html_document

bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
font-family: Times;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  font-family: Times;
}

</style>
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
klippy::klippy(color = 'darkred')
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# intro second draft :
# Introduction

## Research Problem and Goal

In this study, I examine the role of gender homophily in shaping co-authorship networks in the Sociology and Political Science Departments at Radboud University. Gender homophily refers to the tendency of individuals to collaborate with others of the same gender. This phenomenon can influence who gains access to collaboration opportunities, thereby shaping academic trajectories and reinforcing existing inequalities. Understanding these dynamics is essential to uncovering how collaboration networks may perpetuate or mitigate gender-based disparities in academia.

### Gender and Collaboration

Previous research has consistently shown that gender plays a significant role in academic collaboration. Studies by Horta (2022) and Kwiek and Roszka (2021) reveal that male researchers are more likely to collaborate with other men, while female researchers exhibit weaker but still present homophilous tendencies. This pattern contributes to the persistence of gender imbalances in co-authorship, where all-male teams remain more common than all-female teams. These trends are particularly concerning in fields such as Sociology and Political Science, where collaboration is vital for career advancement.

The structure of these networks affects not only who collaborates but also how academic recognition and resources are distributed. For example, male-dominated networks may facilitate better access to high-impact collaborations and grant opportunities for men, further perpetuating the gender gap in academic success.

### Prestige and Collaboration

Another crucial factor influencing collaboration is academic prestige, often measured by publication in high-impact journals. Scholars with higher prestige, indicated by metrics such as the h-index or publication in top-tier journals, are generally more sought after as collaborators. Existing research suggests that prestige might interact with gender homophily in significant ways. Bellotti et al. (2022) found that prestigious male scholars often form networks that exclude women, while McDowell, Singell, and Stater (2006) observed that female scholars, despite demonstrating similar academic competence, face greater challenges in accessing high-prestige collaborative networks.

In this context, it is crucial to explore whether the tendency for gender-based collaboration shifts when prestige factors come into play. Specifically, does publishing in prestigious journals mitigate or exacerbate gender homophily in academic networks?

## Research Focus and Questions

This study aims to analyze gender homophily within the co-authorship networks of the Sociology and Political Science Departments at Radboud University, focusing on the role of journal prestige. The research is guided by the following questions:

- **RQ1a:** To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?

- **RQ1b:** To what extent do the two departments differ in their levels of gender homophily?

These questions focus on descriptive patterns, providing a macro-level view of gender clustering in academic networks.

- **RQ2:** To what extent can gender homophily in collaboration networks be explained by preferences at the dyadic level, controlling for structural network and individual-level covariates?

This question shifts the focus to the micro-level, examining whether individual choices and structural factors drive homophily in co-authorship.

- **RQ3:** How does the prestige of scholars, measured by their publication in high- versus low-impact journals, influence gender homophily in co-authorship networks?

This question explores whether prestige interacts with gender homophily, shedding light on how academic recognition impacts collaborative behaviors. Specifically, it investigates whether male and female scholars with similar prestige levels experience different patterns of collaboration and whether high-prestige contexts reduce or amplify gender clustering.

## Relevance and Contribution

This study contributes to the growing body of literature on gender inequalities in academia by focusing on the intersection of gender and academic prestige in shaping collaboration networks. By applying a social network perspective, it offers insights into how gender-based preferences influence co-authorship patterns and how these dynamics vary between disciplines. Moreover, by exploring the role of journal prestige, this research highlights how structural factors may intersect with social biases to shape academic opportunities, providing a nuanced understanding of the mechanisms underlying gender disparities in academia.

---

# Introduction draft

## Contextualization
Collaboration has become a cornerstone of modern academic practice, with scholars increasingly working together to leverage diverse expertise, share resources, and enhance their research output. Co-authorship, in particular, is a key indicator of collaboration, often reflecting the dynamics of professional relationships and the distribution of academic opportunities. As Cheong (@Cheong2009) highlights, collaboration enables researchers to combine intellectual and institutional resources, contributing significantly to their scholarly success.

However, collaboration is not free from social patterns and biases. Academic networks often mirror broader societal inequalities, with persistent disparities in opportunities and recognition. One such disparity is gender inequality, which continues to shape interactions, opportunities, and outcomes within academia.

## Research problem & goal

Understanding how gender affects co-authorship networks, particularly in terms of homophily, can reveal important insights into how academic opportunities and achievements are distributed across genders.
## N: Research Problem and Goal
In this study, I examine the role of gender homophily in shaping co-authorship networks in the Sociology and Political Science Departments at Radboud University. Gender homophily refers to the tendency of individuals to collaborate with others of the same gender. This phenomenon can influence who gains access to collaboration opportunities, thereby shaping academic trajectories and reinforcing existing inequalities. Understanding these dynamics is essential to uncovering how collaboration networks may perpetuate or mitigate gender-based disparities in academia.

## Short overview of prior findings:

### gender & collaboration

Previous studies have found that male researchers often prefer collaborating with other men, while female researchers tend to collaborate with other women, albeit to a lesser extent (@Horta2022). This raises critical questions about the structure and inclusiveness of academic networks. In fields such as Sociology and Political Science, where collaboration is crucial for scholarly success, gender-based patterns could shape the career trajectories of researchers (@Cheong2009). Furthermore, as Kwiek and Roszka (@Kwiek2021) point out, all-male authorship is far more prevalent than all-female authorship, indicating that male-dominated collaboration networks may reinforce existing gender imbalances in academia.
### N:Gender and Collaboration

Previous research has consistently shown that gender plays a significant role in academic collaboration. Studies by Horta (2022) and Kwiek and Roszka (2021) reveal that male researchers are more likely to collaborate with other men, while female researchers exhibit weaker but still present homophilous tendencies. This pattern contributes to the persistence of gender imbalances in co-authorship, where all-male teams remain more common than all-female teams. These trends are particularly concerning in fields such as Sociology and Political Science, where collaboration is vital for career advancement.

The structure of these networks affects not only who collaborates but also how academic recognition and resources are distributed. For example, male-dominated networks may facilitate better access to high-impact collaborations and grant opportunities for men, further perpetuating the gender gap in academic success.

### h-index & collaboration

Another influential factor in collaboration between scholars seems to be the scholars' 'prestige', namely whether they have published in renowned journals. **SOURCE** found that .....PART ABOUT JOURNAL PRESTIGE

In this study, I explore the extent to which gender homophily influences co-authorship networks among faculty in the Sociology and Political Science Departments at Radboud University. Specifically, I aim to examine whether there are differences in gender homophily between the two departments, and how homophily patterns vary depending on the prestige of the journals in which researchers publish.

### Reiteration of (societal) relevance of study

This research will contribute to a growing body of literature that seeks to understand the social dynamics underlying academic collaboration. By analyzing gender homophily and network structures in two different academic disciplines, this study will shed light on how gender shapes collaboration opportunities and how these dynamics may differ across scholarly fields. Moreover, exploring the role of journal prestige in shaping homophily will offer insights into whether gender-based preferences shift in response to career-enhancing factors like publishing in high-prestige journals.

# Written references (remove in final paper)

Bornmann, Lutz, Christian Ganser, and Alexander Tekles. 2021. “Simulation of the h Index Use at University Departments Within the Bibliometrics-Based Heuristics Framework: Can the Indicator Be Used to Compare Individual Researchers?” *Journal of Informetrics* 16 (1): 101237. <https://doi.org/10.1016/j.joi.2021.101237>.

@Bornmann2021

Bornmann, Lutz, Christian Ganser, Alexander Tekles, and Loet Leydesdorff. 2020. “Does the h αα-Index Reinforce the Matthew Effect in Science? The Introduction of Agent-Based Simulations into Scientometrics.” *Quantitative Science Studies* 1 (1): 331–46.

@Bornmann2020

# References