### 2024-11-23 descriptivestry.Rmd
modelestimates-GOFs-desresults-darkgraph

---
title: "Descriptive adventures"
author: "Verooo"
date: "2024-11-08"
output: html_document
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>My Descriptives</title>

<link rel="stylesheet" href="styles.css"> <!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Getting started

```{r clean}
rm(list = ls())
```

## Packages

```{r package, hide=TRUE}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)
require(ggplot2)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

require(RColorBrewer)

```

## Functions

Functions chunk

```{r functions 1, hide =TRUE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2, hide=TRUE}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

## Load data

Scholar data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

```{r saving scholar output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

Ego data

```{r df_ego data}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

```{r matrix data ego}

test_w1 <- igraph::graph_from_adjacency_matrix(
  test$nets[1,,], # first wave of data. 
  # (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL)

```

# Descriptives - Demographics 
The following graphs represent the segregation of the network based on gender and on H-index for the wave of interest (2015-2018).

## Ego level

### Gender

descriptives (mean etc.) in whole sample

```{r sum gender}
summary(df_ego$gender)
```

gender counts across whole sample (0 = female, 1 = male)

```{r counts gender-all}
table(df_ego$gender)
```

Histogram: overall gender distribution in sample

```{r histogram gender }
# Convert gender to a factor to avoid errors
df_ego$gender <- as.factor(df_ego$gender)
# Create the histogram with a dark theme and two colors for bars
ggplot(df_ego, aes(x = gender, fill = gender)) +
  geom_bar(color = "darkgrey", width = 0.6) +  # Add borders for bars
  scale_fill_manual(
    values = c("0" = "#FFFFB3", "1" = "#80B1D3"),  # Custom colors: yellow for females, blue for males
    labels = c("Female", "Male")  # Optional: customize legend labels
  ) +
  labs(
    title = "Gender Distribution",
    x = "Gender\n (0 = female, 1 = male)",
    y = "Count",
    fill = "Gender"  # Legend title
  ) +
  theme_minimal(base_family = "Arial") +  # Clean, minimal base theme
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark panel background
    panel.grid.major = element_line(color = "#444444"),  # Subtle grid lines
    panel.grid.minor = element_line(color = "#444444"),
    axis.text = element_text(color = "#f0f0f0"),  # White axis text
    axis.title = element_text(color = "#f0f0f0", face = "bold"),  # Bold white axis titles
    plot.title = element_text(color = "#f0f0f0", face = "bold", hjust = 0.5),  # White centered title
    legend.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark legend background
    legend.text = element_text(color = "#f0f0f0"),  # White legend text
    legend.title = element_text(color = "#f0f0f0")  # White legend title
    )

```

Genders per department (0 = female, 1 = male)

```{r gender per dep.}

# Wave 1: Gender count per department 
table(df_ego$discipline.24, df_ego$gender)
```

### h index

descriptives (mean etc.)

```{r summary h index}
summary(df_ego$h_index)
```

Overall(2015-2018)

```{r hist h index}
# Histogram of prestige scores for Wave 1

# Create a data frame for ggplot2 
df_h_index <- data.frame(h_index = df_ego$h_index)

# Generate the histogram with ggplot2
ggplot(df_h_index, aes(x = h_index)) +
  geom_histogram(
    bins = 30,
    fill = "#FFFFB3",  # Bar color
    color = "black"    # Border color for bars
  ) +
  labs(
    title = "Distribution of H-Index from 2015 to 2018",
    x = "H Index",
    y = "Frequency"
  ) +
  theme_minimal(base_family = "Arial") +  # Clean, minimal base theme
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark panel background
    panel.grid.major = element_line(color = "#444444"),  # Subtle grid lines
    panel.grid.minor = element_line(color = "#444444"),
    axis.text = element_text(color = "#f0f0f0"),  # White axis text
    axis.title = element_text(color = "#f0f0f0", face = "bold"),  # Bold white axis titles
    plot.title = element_text(color = "#f0f0f0", face = "bold", hjust = 0.5)  # White centered title
  )

```

H index per department

```{r table per dep h index}
# Wave 1: Prestige distribution by dep. and total
df_ego %>%
    group_by(discipline.24) %>%
    summarise(mean_h_index = mean(h_index, na.rm = TRUE), min_h_index = min(h_index,
        na.rm = TRUE), max_h_index = max(h_index, na.rm = TRUE), sd_h_index = sd(h_index,
        na.rm = TRUE)) %>%
    bind_rows(df_ego %>%
        summarise(discipline.24 = "total", mean_h_index = mean(h_index, na.rm = TRUE), min_h_index = min(h_index,
            na.rm = TRUE), max_h_index = max(h_index, na.rm = TRUE), sd_h_index = sd(h_index,
            na.rm = TRUE)))
```