### 2024-11-23 draft1desresults.Rmd

---
title: "Draft Result Descriptives"
author: "Verooo"
date: "2024-11-23"
output: html_document
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>My Descriptives</title>

<link rel="stylesheet" href="styles.css"> <!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Getting started

```{r clean}
rm(list = ls())
```

## Packages

```{r package, hide=TRUE}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)
require(ggplot2)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

require(RColorBrewer)

```

## Functions

Functions chunk

```{r functions 1, hide =TRUE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2, hide=TRUE}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

## Load data

Scholar data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

```{r saving scholar output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

Ego data

```{r df_ego data}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

```{r matrix data ego}

test_w1 <- igraph::graph_from_adjacency_matrix(
  test$nets[1,,], # first wave of data. 
  # (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL)

```

## Network Description

The network data we are looking at pertains to 101 scholars and their publications. In the time frame of the wave at hand, namely form 2015 to 2018, we find 47 edges, with 6 reciprocated collaborations 36 ties that are asymmetric. Additionally, there were 5009 occasions in which a ties could have been made but was not. The edge density of 0.004 suggests that from all possible edges, this sample contains relatively few.

In terms of triad-configurations, pertaining to the different possibilities three scholars could collaborate, we find that the most common one is no collaboration at all (configuration '003'). Beyond that, the configuration '012' is also common, meaning that only one directed edge is present in the triad. In our sample it is also common for such edge to be reciprocated, as seen in the third-highes-t count of the triad-configuration '102', a graph with a mutual connection between two vertices. Overall, the seems to be high number of isolates in the sample network.

### Nr. of nodes/vertices

```{r nodes}
vcount(test_w1)
```

### Nr. of edges

```{r edges}
ecount(test_w1)
```

### Dyad census

```{r dyad cen}
igraph::dyad_census(test_w1)
```

### Triad-census

```{r triad cen}
igraph::triad_census(test_w1)
```

```{=html}
<!-- 
TRIAD CONFIGURATIONS (16 possible)

o   Three digits [x, x, x]
  First: nr. of reciprocal ties 
  Middle counts: nr of directed ties
  On right: if no ties
o   D, U, C 
  Down 
  Up 
  Cyclical

1. # 003 A,B,C, the empty graph.

2. # 012 A->B, C, the graph with a single directed edge.

3. # 102 A<->B, C, the graph with a mutual connection between two vertices.

4. # 021D A<-B->C, the out-star.

5. # 021U A->B<-C, the in-star.

6. # 021C A->B->C, directed line.

7. # 111D A<->B<-C.

8. # 111U A<->B->C.

9. #- 030T A->B<-C, A->C.

10.#X 030C A<-B<-C, A->C.

11.#- 201 A<->B<->C.

12.#X 120D A<-B->C, A<->C.

13.#- 120U A->B<-C, A<->C.

14.#- 120C A->B->C, A<->C.

15.#- 210 A->B<->C, A<->C.

16.#X 300 A<->B<->C, A<->C, the complete graph.

-->
```
### Edge-density

```{r edge dens}
# edge density
igraph::edge_density(test_w1)
```

### more in depth:

````{=html}
<!--

#### reciprocity

```{r recip}
reciprocity(test_w1, ignore.loops = TRUE, mode = c("default", "ratio"))
```

-->
````

#### Degree centrality

Degree centrality, often simply called 'degree', measures the number of connections a scholar, or ego, has with other scholars, known as alters. In the context of our dataset, we distinguish between two types of degrees: incoming connections (indegrees) and outgoing connections (outdegrees). Indegrees represent ties directed toward the ego, which in this case signify instances where the ego is **not** the primary author of a paper. Conversely, outdegrees represent ties originating from the ego, indicating collaborations where the ego assumes the role of the first author.

```{r degree}
# Calculate degree centrality
degree_testw1 <- degree(test_w1)
degree_testw1 <- as.data.frame(degree_testw1)

ggplot(degree_testw1, aes(x = as.numeric(degree_testw1))) +
  geom_histogram(binwidth = 0.9, fill = "#FFFFB3", color = "black") +  
  labs(
    title = "Distribution Degrees Wave 1",
    x = "Distribution of Degrees",
    y = "Count"
  ) +
  ylim(0, 15) +  # Set y-axis limits 
  theme_minimal(base_family = "Arial") +  # Minimal, clean base theme
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark panel background
    panel.grid.major = element_line(color = "#444444"),  # Subtle grid lines
    panel.grid.minor = element_line(color = "#444444"),
    axis.text = element_text(color = "#f0f0f0"),  # White axis text
    axis.title = element_text(color = "#f0f0f0", face = "bold"),  # Bold white axis titles
    plot.title = element_text(color = "#f0f0f0", face = "bold", hjust = 0.5),  # White centered title
    legend.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark legend background
    legend.text = element_text(color = "#f0f0f0")  # White legend text
  ) +
  guides(fill = "none")  # Remove the legend

```

<!-- ylim was 65 at some point priot due to one higher bar, I wonder what changed since... -->

The indegree distribution is right-skewed. This suggests that only very few vertices have many infegrees, and that most vertices have a low number of indegrees.

#### Betweenness Centrality

(with isolates)

```{r betweenness}
# Calculate betweenness centrality
betweenness_w1 <- igraph::betweenness(test_w1, directed = TRUE)
betweenness_w1 <- as.data.frame(betweenness_w1)

ggplot(betweenness_w1, aes(as.numeric(betweenness_w1), colour = as.numeric(betweenness_w1))) +
  geom_histogram(binwidth = 1.2, fill = "#FFFFB3") +
  labs(
    title = "Distribution Betweenness Wave 1",
    x = "Distribution of Betweenness",
    y = "Count") +
   theme_minimal(base_family = "Arial") +  # Minimal, clean base theme
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark panel background
    panel.grid.major = element_line(color = "#444444"),  # Subtle grid lines
    panel.grid.minor = element_line(color = "#444444"),
    axis.text = element_text(color = "#f0f0f0"),  # White axis text
    axis.title = element_text(color = "#f0f0f0", face = "bold"),  # Bold white axis titles
    plot.title = element_text(color = "#f0f0f0", face = "bold", hjust = 0.5),  # White centered title
    legend.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark legend background
    legend.text = element_text(color = "#f0f0f0")  # White legend text
  ) +
  guides(fill = guide_legend(title = "Betweenness (rounded)"))  # Customize legend

```

#### Clustering

```{r}
# Clustering the wave 
clustering_w1 <- igraph::transitivity(test_w1, type = "local")
clustering_w1 <- as.data.frame(clustering_w1)

# Generate the dark mode plot with a single color fill
ggplot(clustering_w1, aes(x = as.numeric(clustering_w1))) +
   geom_histogram(binwidth = 0.05, fill = "#FFFFB3", color = "black", position = "dodge") +  
  labs(
    title = "Distribution Clustering Wave 1",
    x = "Distribution of Clustering",
    y = "Count"
  ) +
  theme_minimal(base_family = "Arial") +  # Clean, minimal base theme
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark panel background
    panel.grid.major = element_line(color = "#444444"),  # Subtle grid lines
    panel.grid.minor = element_line(color = "#444444"),
    axis.text = element_text(color = "#f0f0f0"),  # White axis text
    axis.title = element_text(color = "#f0f0f0", face = "bold"),  # Bold white axis titles
    plot.title = element_text(color = "#f0f0f0", face = "bold", hjust = 0.5),  # White centered title
    legend.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark legend background
    legend.text = element_text(color = "#f0f0f0")  # White legend text
  ) + guides(fill = "none")  # Remove the legend (since all bars are the same color)

```

# Network-level descriptives

## x. Average path length

```{r path length}
mean((ego_size(test_w1, order = 2, mode = "out") - 1)/vcount(test_w1))
```

## STILL DO! Segregation - Mortons I - also graphs below

**STILL DO!**

# Plotting:

## Plots WITH isolates 
###Gender 
#### whole network
```{r plot gender}

plot(test_w1,
     vertex.color = ifelse(df_ego$gender == 1, "#FFED6F", "#80B1D3"), # now, I can use actor attributes for plotting.
# here: if gender = 1 (aka man) then red dot, if not then blue 
  vertex.label = NA,
  vertex.size = 7.5,
  edge.width = 0.5,
  edge.arrow.size = 0.1,
  main = "Network with isolates [yellow = female, blue = male]",)

```

#### per department

```{r with iso gender per dep old code }

indegr_w1 <- degree(test_w1, mode = "in")

# assign gender to scholar
V(test_w1)$gender <- df_ego$gender

# assign discipline to scholar
V(test_w1)$discipline <- df_ego$discipline.24

V(test_w1)$indegree <- indegr_w1

p_gender <- ggraph(test_w1, layout = "fr") + geom_edge_link(end_cap = circle(4, 'mm'),color = "#D9D9D9", width = 0.7, arrow = arrow(length = unit(2, "mm"), type = "closed")) + geom_node_point(aes(size = indegree,
    color = as.factor(discipline), shape = as.factor(gender))) + scale_size(range = c(3, 10)) + theme_void() +  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_gender + labs(title = "Departments with isolates", subtitle = "Plot of indegree and gender", caption = "Gender: 0 = female, 1 = male.\n Data refer to scholars from Radboud University.") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))
```

### H-Index 
#### whole network 
```{r  h ind with iso old code }
indegr_w1 <- degree(test_w1, mode = "in")

# assign h-index to scholar
V(test_w1)$h_index <- df_ego$h_index
V(test_w1)$indegree <- indegr_w1

# set size of plot
options(repr.plot.width = 15, repr.plot.height = 18)

# plot wave 1 with isolates
p_iso <- ggraph(test_w1, layout = "fr") + geom_edge_link(end_cap = circle(4, 'mm'),color = "#D9D9D9", width = 0.3, arrow = arrow(length = unit(1, "mm"), type = "closed")) + geom_node_point(aes(size = indegree,
    color = h_index)) + scale_size(range = c(3, 10)) +
  scale_color_gradient(low = "#F6222E", high = "#3283FE") +  # Retain original gradient
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
    )

p_iso + labs(title = "Network with isolates", subtitle = "Plot of indegree and H-index", caption = "Data refer to: Scholars RU sociology and political science departments") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"), plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 1))
```

### H-Index and Gender
#### whole network 
```{r dark g&h with iso}
indegr_w1 <- degree(test_w1, mode = "in")

#assign gender to scholar 
V(test_w1)$gender <- df_ego$gender
# V(test_w1)$gender <- ifelse(df_ego$gender >= 0.5,1,0)

# assign h-index to scholar
V(test_w1)$h_index <- df_ego$h_index
V(test_w1)$indegree <- indegr_w1

p_gender <- ggraph(test_w1,
  layout = "fr") +
  geom_edge_link(end_cap = circle(4, 'mm'),color = "#D9D9D9", width = 0.3, arrow = arrow(length = unit(1, "mm"), type = "closed")) +  # Darker edge color
  geom_node_point(aes(size = indegree, color = h_index, shape = as.factor(gender))) +   
  scale_size(range = c(3, 10)) + 
  scale_color_gradient(low = "#F6222E", high = "#3283FE") +  # Retain original gradient
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_gender + 
  labs(title = "Network with isolates",
    subtitle = "Plot of indegree, H-index and gender",
    caption = "Gender: 0 = female, 1 = male.\nData refer to: Scholars RU sociology and political science departments.") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))

```

---

## Plots WITHOUT isolates 
The isolates of the network have been removed to allow for a better readability of the graphs. Only vertices with edges remain part of the visualizations. In other words, only scholars with collaborations are seen in the following graphs.

```{r removing isolates}
indegr_w1 <- igraph::degree(test_w1, mode = "in")

isolates_w1 <- V(test_w1)[degree(test_w1, mode = "all") == 0]

noisolates_w1 <- delete_vertices(test_w1, isolates_w1)

gsize(noisolates_w1) # edges
gorder(noisolates_w1) # vertices 
```

### Gender 
#### DESCIP<> whole network

```{r plot wo isolates}
plot(noisolates_w1,
  vertex.color = ifelse(df_ego$gender == 1, "#80B1D3", "#FFED6F"), # now, I can use actor attributes for plotting.
  # here: if gender = 1 (aka man) then red dot, if not then blue 
  vertex.label = NA,
  vertex.size = 9,
  vertex.frame.color = "black",
  edge.curved = 0.1,
  edge.arrow.size = 0.5,
  edge.color = "#999999",
  edge.width = 0.5,
  main = "Network without isolates [yellow = female, blue = male]",
  asp=0)
```
#### RESULTS <> indegree & department

```{r without iso gender per dep old code}
# assign gender to scholar
V(test_w1)$gender <- df_ego$gender

# assign disciplines to scholars
V(test_w1)$discipline <- df_ego$discipline.24
V(test_w1)$indegree <- indegr_w1

# remove isolates
noisolates_w1 <- delete_vertices(test_w1, V(test_w1)[degree(test_w1) == 0])

p_noisgender <- ggraph(noisolates_w1, layout = "fr") + geom_edge_link(end_cap = circle(4, 'mm'),color = "#D9D9D9", width = 0.7, arrow = arrow(length = unit(2, "mm")), type = "closed") +
    geom_node_point(aes(size = indegree, color = as.factor(discipline), shape = as.factor(gender))) +
    scale_size(range = c(3, 10)) + theme_void() +   theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_noisgender + labs(title = "Departments without isolates", subtitle = "Plot of indegree and gender",
    caption = "Gender: 0 = female, 1 = male.\n Data refer to scholars from Radboud University.") + theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))
```

### H-index 
#### RESULTS<> Whole network (indegree)
```{r}
indegr_w1 <- degree(test_w1, mode = "in")

# assign h-index to scholar
V(test_w1)$h_index <- df_ego$h_index
V(test_w1)$indegree <- indegr_w1

# remove isolates
noisolates_w1 <- delete_vertices(test_w1, V(test_w1)[degree(test_w1) == 0])

# set size of plot
options(repr.plot.width = 15, repr.plot.height = 15)

# plot wave 1 without isolates
p_noiso <- ggraph(noisolates_w1, layout = "fr") + geom_edge_link(end_cap = circle(3, 'mm'),color = "#D9D9D9", width = 0.7, arrow = arrow(length = unit(2, "mm"), type = "closed")) + geom_node_point(aes(size = indegree,
    color = h_index)) + scale_size(range = c(3, 10)) + scale_color_gradient(low = "red", high = "blue") +
    theme_void() +   theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_noiso + labs(title = "Network without isolates", subtitle = "Plot of indegree and H-index", caption = "Data refer to: Scholars RU sociology and political science departments") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))
```

#### per department
```{r hind without iso old code}
indegr_w1 <- degree(test_w1, mode = "in")

# assign h-index to scholar
V(test_w1)$h_index <- df_ego$h_index
# assign discipline to scholar
V(test_w1)$discipline <- df_ego$discipline.24

V(test_w1)$indegree <- indegr_w1

# remove isolates
noisolates_w1 <- delete_vertices(test_w1, V(test_w1)[degree(test_w1) == 0])

# set size of plot
options(repr.plot.width = 15, repr.plot.height = 15)

# plot wave 1 without isolates
p_noiso <- ggraph(noisolates_w1, layout = "fr") + geom_edge_link(end_cap = circle(4, 'mm'),color = "#D9D9D9", width = 0.7, arrow = arrow(length = unit(2, "mm"), type = "closed")) + geom_node_point(aes(size = indegree,
    color = h_index, shape = as.factor(discipline))) + scale_size(range = c(3, 10)) + scale_color_gradient(low = "red",
    high = "blue") + theme_void() +   theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_noiso + labs(title = "Departments without isolates", subtitle = "Plot of indegree and H-index", caption = "Data refer to scholars from Radboud University.") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))
```

### RESULTS <> H-index AND Gender (indegree)

```{r dark g&h without iso}
#assign gender to scholar 
V(test_w1)$gender <- df_ego$gender

# assign h-index to scholar
V(test_w1)$h_index <- df_ego$h_index
V(test_w1)$indegree <- indegr_w1

#remove isolates
noisolates_w1 <- delete_vertices(test_w1, V(test_w1)[degree(test_w1) == 0])

p_noisogender <- ggraph(noisolates_w1,
  layout = "fr") +
  geom_edge_link(end_cap = circle(3, 'mm'),color = "#D9D9D9", width = 0.7, arrow = arrow(length = unit(1.75, "mm"), type = "closed")) +  # Darker edge color
  geom_node_point(aes(size = indegree, color = h_index, shape = as.factor(gender))) +   
  scale_size(range = c(3, 10)) + 
  scale_color_gradient(low = "#F6222E", high = "#3283FE") +  # Retain original gradient
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#2d2d2d", color = NA),  # Dark background
    panel.background = element_rect(fill = "#2d2d2d", color = NA),
    text = element_text(color = "#f0f0f0")  # White text
  )

p_noisogender + labs(title = "Network without isolates",
        subtitle = "Plot of indegree, H-index and gender",
        caption = "Gender: 0 = female, 1 = male.\n Data refer to: Scholars RU sociology and political science departments.") +   
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#f0f0f0"), plot.subtitle = element_text(hjust = 0.5, color = "#f0f0f0"), plot.caption = element_text(hjust = 1, color = "#f0f0f0"))

```