### 2024-11-23 estimate_M0.Rmd

---
title: "Estimate_M0"
author: "Verooo"
date: "2024-11-22"
output: html_document
---

<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>My Descriptives</title>
    <link rel="stylesheet" href="styles.css">  <!-- Link to the CSS file I created for dark mode code chunks -->

</head>
</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

NULL MODEL
Testing if the models can be estimated in separate files without error. Easier for me than cleanign the workspace due to oversight of having separate tabs. Of course I still clean the workspace in between but yaknow. Visual distinction.

# Setup
Clean workspace
```{r clean}
rm(list = ls())
```

## Packages

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

Functions chunk

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

GOF auxiliary Functions 
<!--
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.

-->
[unsure if needed. put these here when trying to figure out why sienaGOF function does not work.]
```{r}

GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

# Get Data

## Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

Save the output of the function.

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

```

## Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

## Model 0 - Null Model

**CHANGE** First, I estimate a null model with only reciprocity and
degree. It has a convergence of NUMBER and a Goodness of Fit on the
indegree distribution of 0 (**NUMBER BEOFRE**), indicating that the model
fits the data very badly. This makes sense, given that the model does
not include any network statistics. Both reciprocity and density are
significant for an alpha of NUMBER. It is important to mention that
significance in RSiena means something different than in conventional
statistics. In normal statistics, such as OLS-regression, significance
of estimates refers to the fact that they are not 0 (RELEVANT?). In
RSiena, however, it refers to the fact that the estimates were precisely
determined (Ripley et al. 2011). Going forward with this chapter,
significance will refer to this definition of significance. The density
parameter has a value of NUMBER, which means that people generally do
not want to have a dense network. The reciprocity parameter has a value
of NUMBER, which means that people like to send ties back to people who
have sent a tie to them.

5.1.1 Create SienaData

### Step 1: Define Siena Data & Variables

```{r RSiena define m0}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M0 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

#if there is not a folder called "results" in your repository: make one

```

```{r}
print01Report(mydata_M0, modelname = "./results/results_M0") # create file with report in results folder 
```

### Step 2: Make effects object/structure

```{r RSiena myeff M0}
# define effects structure 
myeff_M0 <- getEffects(mydata_M0)

```

### Step 3: Add Effects

```{r effects M0}
# Null model: just myeff (just degree and reciprocity by default) # model0  #for this when you create Siena Data, do not add the gender variable

myeff_M0 <- includeEffects(myeff_M0, isolateNet, inPop, outAct)

```

### Step 4: Specify model/algorithm

```{r algo M0}

myAlgorithm_M0 <- sienaAlgorithmCreate (projname = "results_M0")

```

### Step 5: Estimate model

```{r estimate M0, eval=FALSE}

# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)

Model0
# maximum convergence >.25

# run again:
Model0.1 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, prevAns= Model0, returnDeps = TRUE)

Model0.1

# fsave(Model0, file ="estimated_M0")
# fsave(Model0.1, file = "estimated_M0.1")

```

```{r load M0 output, echo=FALSE}
Model0 <- fload("./data/processed/estimated_M0_20241123.rda")
Model0

Model0.1 <- fload("./data/processed/estimated_M0.1_20241123.rda")
Model0.1

```

### Step 6: GOF - Goodness of Fit

```{r GOF M0}
GOF_M0.1 <- sienaGOF(Model0.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0.1, main = 'GOF - "Null" Model 0.1 \n Indegree Distribution')
```