### 2024-11-23 estimate_M1.Rmd

---
title: "Estimate_M1"
author: "Verooo"
date: "2024-11-22"
output: html_document
---

<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>My Descriptives</title>
    <link rel="stylesheet" href="styles.css">  <!-- Link to the CSS file I created for dark mode code chunks -->

</head>
</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Setup
Clean workspace. 
```{r clean}
rm(list = ls())
```

## Packages

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

Functions chunk

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

GOF auxiliary Functions 
<!--
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.

-->
[unsure if needed. put these here when trying to figure out why sienaGOF function does not work.]
```{r}

GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

# Get Data

## Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

Save the output of the function.

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

```

MODEL 1
Testing if the models can be estimated in separate files without error. Easier for me than cleanign the workspace due to oversight of having separate tabs. Of course I still clean the workspace in between but yaknow. Visual distinction.

## Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

## Model 1 - Structural Effects

**CHANGE** This model includes the relevant network statistics: InPop,
outAct, isolateNet.

It has a convergence of NUMBER and a Goodness of Fit on the indegree
distribution of NUMBER.

Indegree Popularity (inPop) is defined as the extent to which people who
have a lot of indegrees are more popular to connect with. This parameter
is NUMBER, meaning that actors have a preference to connect with actors
who have a lot of indegrees. The parameter is significant for an alpha
of NUMBER.

Outdegree Activity (outAct) has a value of NUMBER, but is not
significant. IsolateNet is significant andhas a value of NUMBER and
refers to the preference of remaining an isolate. As this value is
positive (IS IT??!), it means that people prefer to be an isolate in the
network.

### Step 1: Define Siena Data & Variables

```{r RSiena define M1}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M1 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

#if there is not a folder called "results" in your repository: make one

```

```{r}
print01Report(mydata_M1, modelname = "./results/results_M1") # create file with report in results folder 
```

### Step 2: Make effects object/structure

```{r RSiena myeff m1}
# define effects structure 
myeff_M1 <- getEffects(mydata_M1)

```

### Step 3: Add Effects

```{r effects m1}
myeff_M1 <- includeEffects(myeff_M1, isolateNet, inPop, outAct)

```

### Step 4: Specify model/algorithm

```{r algo m1}

myAlgorithm_M1 <- sienaAlgorithmCreate (projname = "results_M1")

```

### Step 5: Estimate model

```{r estimate m1}
# estimate the model 
Model1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, returnDeps = TRUE)
#latter needed for GOF

Model1

Model1.1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, prevAns =Model1, returnDeps = TRUE)
#latter needed for GOF 
Model1.1

```

### Step 6: GOF - Goodness of Fit

```{r GOF M1}
GOF_M1 <- sienaGOF(Model1.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M1, main = "GOF - Model 1 \n Indegree Distribution")
```