### 2024-11-23 estimate_M2.Rmd

---
title: "Estimate_M2"
author: "Verooo"
date: "2024-11-22"
output: html_document
---

<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>My Descriptives</title>
    <link rel="stylesheet" href="styles.css">  <!-- Link to the CSS file I created for dark mode code chunks -->

</head>
</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

MODEL 2
Testing if the models can be estimated in separate files without error. Easier for me than cleanign the workspace due to oversight of having separate tabs. Of course I still clean the workspace in between but yaknow. Visual distinction.

# Setup
clean workspace
```{r clean}
rm(list = ls())
```

## Packages

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

Functions chunk

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

GOF auxiliary Functions 
<!--
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.

-->
[unsure if needed. put these here when trying to figure out why sienaGOF function does not work.]
```{r}

GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

# Get Data

## Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

Save the output of the function.

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

```

## Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

## Model 2 - Gender

**CHANGE** Model 2 builds on Model 1 and includes the network statistics
for gender: diffX, diffXTransTrip, egoX & altX. This model has a
convergence of 0.23 and a Goodness of Fit on the indegree distribution
of 0.184. For the structural network statistics we find similar results,
in terms of estimate size and significance, as in Model 1. However, the
outdegree activity parameter almost reaches significance (-1.7135484).
With regard to the covariate network statistics, we find that the gender
of ego (egoX), the gender of alter (altX) and the difference between ego
and alter (diffX) is not significant and thus the estimate cannot be
trusted. The transtrip for different gender (diffXTransTrip) is positive
and significant, which implies that an ego prefers to close triads with
alters who have a different gender score than theirs. This is an
interesting finding, because it implies that their is a tendency towards
diverse collaboration. However, when compared to the level of gender
clustering and the fact that this clustering decreased, this finding
does make sense.

### Step 1: Define Siena Data & Variables

```{r RSiena define m2}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# define data object 
mydata_M2 <- sienaDataCreate(nets, gender_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

#if there is not a folder called "results" in your repository: make one

```

```{r}
print01Report(mydata_M2, modelname = "./results/results_M2") # create file with report in results folder 
```

### Step 2: Make effects object/structure

```{r RSiena myeff m2}
# define effects structure 
myeff_M2 <- getEffects(mydata_M2)

```

### Step 3: Add Effects

```{r effects m2}
myeff_M2 <- includeEffects(myeff_M2, isolateNet, inPop, outAct)

myeff_M2 <- includeEffects(myeff_M2, egoX, altX, sameX, interaction1 = "gender_RS")

```

### Step 4: Specify model/algorithm

```{r algo m2}

myAlgorithm_M2 <- sienaAlgorithmCreate (projname = "results_M2")

```

### Step 5: Estimate model

```{r estimate m2}
# estimate the model 
Model2 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, returnDeps = TRUE)
#latter needed for GOF

Model2

```

### Step 6: GOF - Goodness of Fit

```{r GOF M2}
GOF_M2 <- sienaGOF(Model2, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M2, main = "GOF - Model 2 \n Indegree Distribution")
```