### 2024-11-23 estimate_M3.Rmd

---
title: "Estimate_M3"
author: "Verooo"
date: "2024-11-22"
output: html_document
---

<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>My Descriptives</title>
    <link rel="stylesheet" href="styles.css">  <!-- Link to the CSS file I created for dark mode code chunks -->

</head>
</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

MODEL 3 - Gender & H-index 
Testing if the models can be estimated in separate files without error. Easier for me than cleanign the workspace due to oversight of having separate tabs. Of course I still clean the workspace in between but yaknow. Visual distinction.

# Setup
Clean workspace.
```{r clean}
rm(list = ls())
```

## Packages

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

Functions chunk

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

GOF auxiliary Functions 
<!--
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.

-->
[unsure if needed. put these here when trying to figure out why sienaGOF function does not work.]
```{r}

GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

# Get Data

## Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
#make sure to lead fload first of course
```

Save the output of the function.

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

```

## Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

## Model 3 - H-Index

**CHANGE!!!** Model 3 includes the covariate network statistics for
Prestige: egoX, altX and inPopX. This model has a convergence ratio of
0.23 and a a Goodness of Fit on the indegree distribution of 0.64. In
this model, the parameters for density and reciprocity are still
significant, with values of -1.98 (se = 0.71) and 2.59 (se = 0.38)
respectively, meaning that actors prefer sparse networks, but do prefer
to send ties to alters who have sent ties to them. IsolateNet is still
significant with a value of 3.23 (se = 1.06), implying that actors
prefer to be isolates. In this model, indegree popularity squared (b =
0.17, se = 0.21) is no longer significant, implying that this might be
explained by other network statistics in the model. The covariate
statistics for gender show similar findings as Model 2 & Model 2.5,
namely that actors prefer to close triads with alters that have a
different score for gender (b = 1.51, se = 0.44). the ego effect and
alter effect for gender are non-significant, as was the case in Model 2
& Model 2.5. Lastly, Model 3 shows some significant results with regards
to the covariate network statistics for prestige. The prestige effect
for alters is positive, implying that people prefer to send ties to
people who have a higher amount of prestige (b = 0.042, se = 0.0089).
Additionally, the prestige effect for egos is negative, which indicates
that people who have higher prestige send less ties (b = -0.030, se =
0.015). This is in line with theories of preferential attachment and
upward mobility, because as egos with higher prestige will are more
likely to receive request for collaboration rather then sending request
for collaboration. Interestingly enough, the indegree popularity
statistic for prestige is not significant (b = -0.012, se = 0.023),
implying that actors who are more prestigious do not receive more
indegrees.

### Step 1: Define Siena Data & Variables

```{r RSiena define m3}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index

# covariate 2
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata_M3 <- sienaDataCreate(nets, gender_RS, h_index_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)
# if there is not a folder called "results" in your repository: make one

```

```{r}
print01Report(mydata_M3, modelname = "./results/results_M3") # create file with report in results folder 
```

### Step 2: Make effects object/structure

```{r RSiena myeff M3}

# define effects structure 
myeff_M3 <- getEffects(mydata_M3)

```

### Step 3: Add Effects

```{r effects M3}
myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model1

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, sameX, interaction1 = "gender_RS") # model2

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, inPopX, interaction1 = "h_index_RS") # model3

```
<!-- # isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collab with people who are popular

# change statistics to fit whatever I want to know about h-index 
# not sameX 
# look up inPopX: preferential attachment (cf. Koen lel), higher h-index more popular 
  # cf. Fridas modeling process (Jochem told her that )
-->
### Step 4: Specify model/algorithm

```{r algo M3}

myAlgorithm_M3 <- sienaAlgorithmCreate (projname = "results_M3")

```

### Step 5: Estimate model

```{r estimate M3}
# estimate the model 
Model3 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, returnDeps = TRUE)
#latter needed for GOF

Model3

```

### Step 6: GOF - Goodness of Fit

```{r GOF M3}
GOF_M3 <- sienaGOF(Model3, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3, main = "GOF - Model 3 \n Indegree Distribution")
```