### 2024-11-23 finalGOF.Rmd

---
title: "GOF"
author: "Verooo"
date: "2024-11-23"
output: html_document
---