### 2024-11-23 finalestimate.Rmd

---
title: "Model Estimation"
author: "Verooo"
date: "2024-11-23"
output: html_document
---