### 2024-11-23 finalintro.Rmd

---
title: "Introduction"
author: "Verooo"
date: "2024-11-23"
output: html_document
---