### 2024-11-23 modelestimate.Rmd

---
title: "Model estimation"
author: "Verooo"
date: "2024-11-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>My Descriptives</title>

<link rel="stylesheet" href="styles.css">
<!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Setup

Clean workspace. As always.

```{r clean}
rm(list = ls())
```

## Packages

```{r package, include=TRUE, hide=TRUE, eval=TRUE}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

Functions chunk:

```{r functions 1, hide=TRUE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

More functions:

```{r functions 2, hide= TRUE}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

# Get Data

Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
```

Save the output of the function.

```{r saving output}
# save the output of your function
test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

------------------------------------------------------------------------

# Analysis

```{=html}
<!--
## TEST MODEL

### test 1: Define Siena Data & Variables
-->
```
```{r RSiena define, hide =TRUE, echo=FALSE, eval=FALSE}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index

# covariate 2
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata <- sienaDataCreate(nets, gender_RS, h_index_RS)

```

<!-- Print short descriptive report-->

```{r, hide =TRUE, echo=FALSE, eval=FALSE}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

#if there is not a folder called "results" in your repository: make one

```

```{r, hide =TRUE, echo=FALSE, eval=FALSE}
print01Report(mydata, modelname = "./results/try") # create file with report in results folder 
```

<!-- ### test 2: Make effects object/structure -->

```{r RSiena myeff, hide =TRUE, echo=FALSE, eval=FALSE}

# define effects structure 
myeff <- getEffects(mydata)

```

```{=html}
<!--

e.g. outdegree (density), called density, three things: evaluation,
endowment, and creation you can only include two of the three (tie
creation, breaking and maintaining)

The assumption is: mechanisms to make and break a a tie are the same so
WE ONLY USE THE EVALUATION part

All these effects are described in the manual of RSiena Chapter 12!

-->
```
```{=html}
<!--  
### Step X: Initial description
-->
```
```{=html}
<!--
Has a lot of information in there Also: how tie CHANGED Also: Jaccard
index (measure of stability) Also rate parameter and outdegree

<https://www.stats.ox.ac.uk/~snijders/siena/RSiena_Manual.pdf>

In the package RsienaTwoStep there are functions for the following
network statistics ***s*** :

degree: `ts_degree()`

reciprocity: `ts_recip()`

outdegree activity: `ts_outAct()`

indegree activity: `ts_inAct()`

outdegree popularity: `ts_outPop()`

indegree popularity: `ts_inPop()`

transitivity: `ts_transTrip()`

mediated transitivity: `ts_transMedTrip()`

transitive reciprocated triplets: `ts_transRecTrip()`

number of three-cycles: `ts_cycle3()`
-->
```
<!-- ### test 3: Add Effects -->

```{r effects, hide =TRUE, echo=FALSE, eval=FALSE}

# Null model: just myeff (just degree and reciprocity by default) # model0  #for this when you create Siena Data, do not add the gender variable

myeff <- includeEffects(myeff, isolateNet, inPop, outAct) 
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collab with people who are popular
  # over time: inpop: ppl with lots of indegrees will havelot of indegrees

#n-degree related popularity effect (inPop) (earlier called popularity or popularity of alter effect), defined by the sum of the in-degrees of the others to whom i is tied,

# later (in population) outAct: ppl with lots of outdegrees will have lot
# of indegrees later

myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1 = "gender_RS") # model1
# include effect of h index later too etc
# look up which effects to include

myeff <- includeEffects(myeff, egoX, altX, inPopX, interaction1 = "h_index_RS") # model 2 (as long as myeff is eithin the brackets too, the other things get added to the myeff object)

# change statistics to fit whatever I want to know about h-index 
# not sameX 
# look up inPopX: preferential attachment (cf. Koen lel), higher h-index more popular 
  # cf. Fridas modeling process (Jochem told her that )

#diffXTransTrip
#different covariate × transitive triplets (diffXTransTrip), defined by the number of transitive triplets i → h → j ← i that have different covariate values for i and j, diff. cov. values = diff gender -RSiena manual

# "The transtrip for different gender (diffXTransTrip) is positive and significant, which implies that an ego prefers to close triads with alters who have a different gender score than theirs. This is an interesting finding, because it implies that their is a tendency towards diverse collaboration. However, when compared to the level of gender clustering and the fact that this clustering decreased, this finding does make sense." - Niels

```

```{=html}
<!--

hint FOR MY OWN PROJECT: interaction effect of these statistics and the
effect of gender

THE "X" in altX and eogX refer to the "interaction 1" variable, so it's eg. ego x gender_RS and alter x gender_RS!

altX - egos send more ties who have more on gender (aka. gender =1) 
(iiiish look at egoX explanantion, this is basically the same but in "reverse", aka about receiving rather than sending)

-   covariate-alter or covariate-related popularity (altX), defined by
    the sum of the covariate over all actors to whom ***i*** has a tie,
-   same as egoX but RECEIVING ties

egoX - [definition?]

-   covariate-ego or covariate-related activity (egoX), defined by i’s
    out-degree weighted by his covariate value,
-   since male = 1 egoX only "looks at" men's ties 
-   if egoX significant then men's ties sign. diff. based on whether its to a man or woman (look at direction of effect to know which gender)

"interaction1" HAS NOTHING TO DO WITH AN INTERACTION EFFECT

-   If there are several dependent variables (networks and/or behavioral
    variables), the name of the dependent variable (name) also is
    required to specify the effect, and will be given in the column
    interaction1 (etc.) of the effects object.

-   It should be mentioned that the keywords interaction1, interaction2,
    and interaction3 in functions such as includeEffects do not point to
    interactions between effects. These keywords are used only to
    exactly specify an effect. They point to the covariate or other
    dependent variable used in the effect; this name was chosen for the
    keyword because this other variable may be said to be interacting
    with the dependent variable (network or behavior).

-->
```
<!-- ### test 4: Specify model/algorithm -->

```{r algo, hide =TRUE, echo=FALSE, eval=FALSE}
# Algorithm from when Jochem and I sat together but it's not fitting anymore bc model type is for undirected ties 
# myAlgorithm <- sienaAlgorithmCreate(modelType = c(nets=6), projname = "proProject")
# undirected: model 6 "additive model", #check manual for this NOW IT'S UNDIRECTED!!!

# Niels advice: take his algorithm instead. haha ok, no questions asked my guy, you got it. 
myAlgorithm_test <- sienaAlgorithmCreate (projname = "results_test")

```

<!-- ### test 5: Estimate model -->

```{r estimate, hide =TRUE, echo=FALSE, eval=FALSE}
# estimate the model 
model_test <- siena07(myAlgorithm_test, data = mydata, effects = myeff, returnDeps = TRUE)
#latter needed for GOF lateron

model_test

```

```{=html}
<!-- 
output:  convergence ration NA probably bc we chose all ties to be included (type = "all") earlier
we can go back and change that or change it here in the myAlgorithm function

convergence ratio should be <.25   !!!
- if above, run model again
- if stays above .25, remove some of the netw. stats
- if close to threshold (e.g. 0.2584) add prevAns = model1  to the siena07 command and rerun then
      model1 <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns= model1, returnDeps = TRUE)

-->
```
<!-- ### test 6: GOF - Goodness of Fit -->

```{r GOF test, hide =TRUE, echo=FALSE, eval=FALSE}
GOF <- sienaGOF(model_test, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF, main = "Test Indegree Distr")

# "Error using packet 1 "bw" is not positive"

#--------- Try to fix error:  -------  
##--------- Inspect GOF object  ------- 
# str(GOF)
  # summary(GOF$Joint$SimulatedTestStat)

##--------- Check for invalid /infinte values ------- 
    # anyNA(GOF$Joint$SimulatedTestStat)  # Should return FALSE
    # any(!is.finite(GOF$Joint$SimulatedTestStat))  # Should return FALSE
    # summary(GOF$Joint$SimulatedTestStat)  # Inspect the range of values

    # Get rid of invalid or infinite values
        # GOF$Joint$SimulatedTestStat <- na.omit(GOF$Joint$SimulatedTestStat)
        # GOF$Joint$SimulatedTestStat <-   GOF$Joint$SimulatedTestStat[is.finite(GOF$Joint$SimulatedTestStat)]

  # var(GOF$Joint$SimulatedTestStat)  # Should return a positive value

##--------- Density simulation  ------- 
    # simulated <- GOF$Joint$SimulatedTestStat
    # # Try density estimation with default bandwidth
    #   density_est <- tryCatch(density(simulated), error = function(e) e)
    #   if (inherits(density_est, "error")) {
    #   print("Density estimation failed.")
    #     } else {
    #             print("Density estimation succeeded.")
    #             plot(density_est, main = "Simulated Test Statistics")
    #     }
    #   
##--------- Force bandwidth -------     
      # Force the Bandwidth in plot.sienaGOF
      # plot(GOF, main = "Indegree Distribution Goodness of Fit", bw = 0.5)
      # #nope does not work

##--------- Custom function -------  
      # custom_plot_sienaGOF <- function(gof, bw = 0.5, ...) {
      # simulated <- gof$Joint$SimulatedTestStat
      # observed <- gof$Joint$ObservedTestStat
      # 
      # # Perform density estimation with forced bandwidth
      # density_est <- density(simulated, bw = bw)
      # 
      # # Plot the density
      # plot(density_est, main = "Custom Indegree Distr.- GOF", ...)
      # abline(v = observed, col = "red", lty = 2, lwd = 2) 
      # }
      # 
      #  #run custom function      
      # custom_plot_sienaGOF(GOF, bw = 0.5)
      # # works but not the same at all 
      # 
##--------- Ggplot2 ------- 
# Fuck this shit. I using ggplot 
    # library(ggplot2)
    # 
    # # Prepare data
    # data <- data.frame(
    #   value = GOF$Joint$SimulatedTestStat,
    #   type = "Simulated"
    # )
    # data <- rbind(data, data.frame(value = GOF$Joint$ObservedTestStat, type = "Observed"))
    # 
    # # Plot density
    # ggplot(data, aes(x = value, fill = type)) +
    #   geom_density(alpha = 0.5, adjust = 1, bw = 0.5) +
    #   geom_vline(xintercept = GOF$Joint$ObservedTestStat, color = "red", linetype = "dashed") +
    #   ggtitle("Indegree Distribution Goodness of Fit") +
    #   theme_minimal()

# works but it's the same thing as with the custom function... It not a violin-plot and just does not hit the spot. 
# The only violins I have are the ones playing the sad self-pity music in my brain right now.

##--------- reinstall RSiena -------

# I.... 
    # ....
      # .... 
        # .... sure.

# --------- I hate myself and back we go to GOF -----
# other siena GOF functions (outdegree does not work either, GeodesicDistribution weirdly enough works.)

#gofo0 <- sienaGOF(model_test, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20), varName = "nets")
#plot(gofo0, main ="Test Outdegree Distr")

#gof0.gd <- sienaGOF(model_test, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "nets")
#plot(gof0.gd, main ="Test Geodesic")

```

```{=html}
<!--

Make it a directed network:  ties from "all" to "first"
focus: writing,
model interpretation

### Questions

-   Is the use of RSiena different for people who look at 1 vs 2 waves.
    I need to look at one, made 2 anyways bc RSiena needs waves to do
    it's thing but:\
    Doesn't it look at ministeps over time? How can I interpret the
    output in a way that pertains to only one wave when the process
    seems to consider it within the construct of 'future change'?

    -   How does my model interpretation potentially differ from peeps
        who looked at two waves (or more)?

-   not looking at TWoStep! diff. modelling process. 
RSiena works with ministeps indeed not just Twosteps (that works with two steps lol)

- diff. betweeen wave 1 and wave 2 model -> convergence 
- default recip. and degree: estimates the putput based on a certain reciproc. and degree "assumption"

-->
```
## \<\> Model 0 - Null Model

First, I estimate a null model with only reciprocity and degree. The
maximum convergence of the model is **NUMBER** The GOF on the Indegree
Distribution is **NUMBER** To have all GOF plots compiled in one place,
see here: *Link GOF page* This indicates that the model is a bad fit for
the data, which was to be expected since we ran a null model without any
network effects.

**CHANGE** Both reciprocity and density are significant for an alpha of
**NUMBER**.

**meaning of significance in RSiena** "It is important to mention that
significance in RSiena means something different than in conventional
statistics. In normal statistics, such as OLS-regression, significance
of estimates refers to the fact that they are not 0 (**RELEVANT?**). In
RSiena, however, it refers to the fact that the estimates were precisely
determined (Ripley et al. 2011). Going forward with this chapter,
significance will refer to this definition of significance." - Niels
Vullings

This model has a density parameter of **NUMBER**, suggesting that
"people generally do not want to have a dense network".

The reciprocity parameter of **NUMBER** indicates that "people like to
send ties back to people who have sent a tie to them."

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m0}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M0 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M0, modelname = "./results/results_M0") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M0}
# define effects structure 
myeff_M0 <- getEffects(mydata_M0)

```

### ---Step 3: Add Effects

```{r effects M0}

myeff_M0 <- includeEffects(myeff_M0, isolateNet, inPop, outAct)

```

<!-- # Null model: just myeff (just degree and reciprocity by default) # model0  #for this when you create Siena Data, do not add the gender variable -->

### ---Step 4: Specify model/algorithm

```{r algo M0}

myAlgorithm_M0 <- sienaAlgorithmCreate (projname = "results_M0")

```

### ---Step 5: Estimate model

```{r estimate M0, eval=FALSE}
# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)
Model0
# maximum convergence >.25
# run again:
Model0.1 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, prevAns= Model0, returnDeps = TRUE)
Model0.1

#fsave(Model0, file ="estimated_M0")
#fsave(Model0.1, file = "estimated_M0.1")

```

```{r load M0 output, echo=FALSE}
Model0 <- fload("./data/processed/estimated_M0_20241123.rda")
Model0

Model0.1 <- fload("./data/processed/estimated_M0.1_20241123.rda")
Model0.1

```

```{r GOF M0}
GOF_M0.1 <- sienaGOF(Model0.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0.1, main = 'GOF - "Null" Model 0 \n Indegree Distribution')
```

------------------------------------------------------------------------

## \<\> Model 1 - Structural Effects

This model additionally entails InPop, outAct, isolateNet as network
statistics.

**CHANGE** "It has a convergence of **NUMBER** and a Goodness of Fit on
the indegree distribution of **NUMBER**."

"Indegree Popularity (inPop) is defined as the extent to which people
who have a lot of indegrees are more popular to connect with. This
parameter is NUMBER, meaning that actors have a preference to connect
with actors who have a lot of indegrees. The parameter is significant
for an alpha of **NUMBER**."

"Outdegree Activity (outAct) has a value of **NUMBER**, but is not
significant. IsolateNet is significant andhas a value of NUMBER and
refers to the preference of remaining an isolate. As this value is
positive (\*IS IT??!\*\*), it means that people prefer to be an isolate
in the network."

### ---Step 1: Define Siena Data & Variables

```{r RSiena define M1}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M1 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M1, modelname = "./results/results_M1") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m1}
# define effects structure 
myeff_M1 <- getEffects(mydata_M1)

```

### ---Step 3: Add Effects

```{r effects m1}

myeff_M1 <- includeEffects(myeff_M1, isolateNet, inPop, outAct)

```

### ---Step 4: Specify model/algorithm

```{r algo m1}

myAlgorithm_M1 <- sienaAlgorithmCreate (projname = "results_M1")

```

### ---Step 5: Estimate model

```{r estimate m1, eval = FALSE}
# estimate the model 
Model1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, returnDeps = TRUE)
Model1
# maximum convergence >.25
# run again:
Model1.1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, prevAns= Model1, returnDeps = TRUE)
Model1.1

#fsave(Model1, file ="estimated_M1")
#fsave(Model1.1, file = "estimated_M1.1")

```

```{r load M1 output, echo=FALSE}

Model1 <- fload("./data/processed/estimated_M1_20241123.rda")
Model1

Model1.1 <- fload("./data/processed/estimated_M1.1_20241123.rda")
Model1.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M1}
GOF_M1.1 <- sienaGOF(Model1.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M1.1, main = "GOF - Model 1 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 2 - Gender

This model is partly similar to Model 1 but additionally holds the
network statistics sameX, diffXTransTrip, egoX & altX of the covariate
gender.

**CHANGE** "This model has a convergence of **0.23** and a Goodness of
Fit on the indegree distribution of **0.184**. For the structural
network statistics we find similar results, in terms of estimate size
and significance, as in Model 1. However, the outdegree activity
parameter almost reaches significance (**-1.7135484**).

With regard to the covariate network statistics, we find that the gender
of ego (egoX), the gender of alter (altX) **are not significant and thus
the estimate cannot be trusted**."

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m2}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# define data object 
mydata_M2 <- sienaDataCreate(nets, gender_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M2, modelname = "./results/results_M2") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m2}
# define effects structure 
myeff_M2 <- getEffects(mydata_M2)
```

### ---Step 3: Add Effects

```{r effects m2}
myeff_M2 <- includeEffects(myeff_M2, isolateNet, inPop, outAct)

myeff_M2 <- includeEffects(myeff_M2, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") 
```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collaborate with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;

#diffXTransTrip
different covariate × transitive triplets (diffXTransTrip), defined by the number
of transitive triplets i → h → j ← i that have different covariate values for i and j,
diff. cov. values = diff gender

-->
```
### ---Step 4: Specify model/algorithm

```{r algo m2}

myAlgorithm_M2 <- sienaAlgorithmCreate (projname = "results_M2")

```

### ---Step 5: Estimate model

```{r estimate m2, eval =FALSE}
# estimate the model 
Model2 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, returnDeps = TRUE)
Model2
# maximum convergence >.25
# run again:
Model2.1 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, prevAns= Model2, returnDeps = TRUE)
Model2.1

#fsave(Model2, file ="estimated_M2")
#fsave(Model2.1, file = "estimated_M2.1")

```

```{r load M2 output, echo=FALSE}

Model2 <- fload("./data/processed/estimated_M2_20241123.rda")
Model2

Model2.1 <- fload ("./data/processed/estimated_M2.1_20241123.rda")
Model2.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M2}
GOF_M2.1 <- sienaGOF(Model2.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M2.1, main = "GOF - Model 2 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 3 - Department

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
test if an alter's department affects an ego's tendency to initiate
collaborations and to see if department affects an individual’s
attractiveness as a collaborator.

**CHANGE** "This model has a convergence of **0.23** and a Goodness of
Fit on the indegree distribution of **0.184**. For the structural
network statistics we find similar results, in terms of estimate size
and significance, as in Model 1. However, the outdegree activity
parameter almost reaches significance (**-1.7135484**).

With regard to the covariate network statistics, we find that the gender
of ego (egoX), the gender of alter (altX) **are not significant and thus
the estimate cannot be trusted**."

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m3}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# define data object 
mydata_M3 <- sienaDataCreate(nets, gender_RS, departm_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M3, modelname = "./results/results_M3") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m3}
# define effects structure 
myeff_M3 <- getEffects(mydata_M3)
```

### ---Step 3: Add Effects

```{r effects m3}
myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model 1

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collaborate with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;

#diffXTransTrip
different covariate × transitive triplets (diffXTransTrip), defined by the number
of transitive triplets i → h → j ← i that have different covariate values for i and j,
diff. cov. values = diff gender

-->
```
### ---Step 4: Specify model/algorithm

```{r algo m3}

myAlgorithm_M3 <- sienaAlgorithmCreate (projname = "results_M3")

```

### ---Step 5: Estimate model

```{r estimate m3, eval =FALSE}
# estimate the model 
Model3 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, returnDeps = TRUE)
Model3

# if maximum convergence >.25: run again
#Model3.1 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, prevAns= Model3, returnDeps = TRUE)
#Model3.1

#fsave(Model3, file ="estimated_M3")
#fsave(Model3.1, file = "estimated_M3.1")

```

```{r load M3 output, echo=FALSE}

Model3 <- fload("./data/processed/estimated_M3_20241123.rda")
Model3

#Model3.1 <- fload ("./data/processed/estimated_M3.1_20241123.rda")
#Model3.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M3}
GOF_M3 <- sienaGOF(Model3, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3, main = "GOF - Model 3 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 4 - H-Index

This model further builds on the prior model and also includes the
network statistics of the covariat H-index, namely **egoX, altX and
inPopX**.

**CHANGE!!!** "This model has a convergence ratio of **0.23** and a a
Goodness of Fit on the indegree distribution of **0.64**. In this model,
the parameters for density and reciprocity are still significant, with
values of **-1.98** (se = **0.71**) and **2.59** (se = **0.38**)
respectively, meaning that actors prefer sparse networks, but do prefer
to send ties to alters who have sent ties to them.

IsolateNet is still significant with a value of **3.23** (se =
**1.06**), implying that actors prefer to be isolates.

In this model, **indegree popularity squared** (b = **0.17**, se =
**0.21**) is no longer significant, implying that this might be
explained by other network statistics in the model.

The covariate statistics for gender show similar findings as Model 2,
namely that actors prefer to close triads with alters that have a
different score for gender (b = **1.51**, se = **0.44**).

The ego effect and alter effect for gender are **non-significant**, as
was the case in Model 2.

Lastly, Model 3 shows some significant results with regards to the
covariate network statistics for prestige. The prestige effect for
alters is positive, implying that people prefer to send ties to people
who have a higher amount of prestige (b = **0.042**, se = **0.0089**).

Additionally, the prestige effect for egos is negative, which indicates
that people who have higher prestige send less ties (b = **-0.030**, se
= **0.015**). **This is in line with theories of preferential attachment
and upward mobility, because as egos with higher prestige will are more
likely to receive request for collaboration rather then sending request
for collaboration.**

Interestingly enough, the indegree popularity statistic for prestige is
not significant (b = **-0.012**, se = **0.023**), implying that actors
who are more prestigious do not receive more indegrees."

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m4}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender
# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index
# covariate 3
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata_M4 <- sienaDataCreate(nets, gender_RS, departm_RS, h_index_RS)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M4, modelname = "./results/results_M4") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M4}

# define effects structure 
myeff_M4 <- getEffects(mydata_M4)

```

### ---Step 3: Add Effects

```{r effects M4}

myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model 1

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, inPopX, interaction1 = "h_index_RS") # model 4

```

```{=html}
<!--
# isolatenet: if individual is alone they "like to/tend to stay alone"
# inPop: ppl like to collab with people who are popular

# sameX
# same covariate, which can also be called covariate-related identity (sameX), defined by the number of ties of i to all other actors j who have exactly the same value on the covariate, s net i95(x) = P jxij I{vi = vj},
# where the indicator function I{vi = vj} is 1 if the condition {vi = vj} is satisfied, and 0 if it is not;

# look up inPopX: preferential attachment (cf. Koen lel), higher h-index more popular 
  # cf. Fridas modeling process (Jochem told her that )

-->
```
### ---Step 4: Specify model/algorithm

```{r algo M4}

myAlgorithm_M4 <- sienaAlgorithmCreate (projname = "results_M4")

```

### ---Step 5: Estimate model

```{r estimate M4, eval =FALSE}
# estimate the model 
Model4 <- siena07(myAlgorithm_M4, data = mydata_M4, effects = myeff_M4, returnDeps = TRUE)
Model4

fsave(Model4, file ="estimated_M4")

```

```{r load M4 output, echo=FALSE}

Model4 <- fload("./data/processed/estimated_M4_20241123.rda")
Model4

```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M4}
GOF_M4 <- sienaGOF(Model4, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M4, main = "GOF - Model 4 \n Indegree Distribution")
```