### 2024-11-24 draftGOF.Rmd
draft discussion

---
title: "GOFdraft"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>GOF</title>

<link rel="stylesheet" href="styles.css"> <!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Setup

Clean workspace.

```{r clean, hide =TRUE}
rm(list = ls())
```

## Packages

```{r package, hide =TRUE}
require(RSiena)

```

## Functions

Functions chunk

```{r functions 1, hide = TRUE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

```

# Goodness of Fit

The goodness of fit (GOF) for a model assesses how well the estimated model fits the observed data. This evaluation is crucial because `RSiena` relies on simulations to estimate changes within the network (@RSiena_manual). GOF measures indicate whether the simulated network matches the changes in the observed network. The use of the specific GOF measure of a model's Indegree Distribution seems most appropriate since the study's focus lies on changes in directed ties between schoalrs. A GOF with a p-value below .05 indicates a poor fit and calls for adjustment of the model, while a p-value above .05 usually indicates an acceptable or good fit. \
The GOF measures of Indegree Distribution for the null model (*p* = .008) and model 1 *(p* = .002) both display p-values below .05. This indicates that they do not fit the data well. This makes sense, since the null model and model 1 do not contain any network effects yet. This also implies that we should not interpret the estimates of these models. The GOF of model 2 (*p* = .096) indicates an acceptable fit, while model 3 (*p* = .12) seems to be even better fitting. Overall, the final model, model 4 (*p* = .197), including all the covariates investigated in this study, has the best fit.

```{=html}
<!--  M0: 0.008  M1: 0.002  M2: 0.096  M3: 0.12  M4: 0.197  -->
```
# Model 0

```{r}
Model0.1 <- fload("./data/processed/estimated_M0.1_20241123.rda")
Model0.1
```

```{r GOF M0}
GOF_M0.1 <- sienaGOF(Model0.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0.1, main = 'GOF - "Null" Model 0.1 \n Indegree Distribution')

```

# Model 1

```{r}
Model1.1 <- fload("./data/processed/estimated_M1.1_20241123.rda")

```

```{r GOF M1}
GOF_M1.1 <- sienaGOF(Model1.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M1.1, main = "GOF - Model 1 \n Indegree Distribution")
```

# Model 2

```{r}
Model2.1 <- fload ("./data/processed/estimated_M2.1_20241123.rda")
Model2.1
```

```{r GOF M2}
GOF_M2.1 <- sienaGOF(Model2.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M2.1, main = "GOF - Model 2 \n Indegree Distribution")
```

# Model 3

```{r, hide=TRUE}
Model3 <- fload("./data/processed/estimated_M3_20241123.rda")
```

```{r GOF M3, hide=TRUE}
GOF_M3 <- sienaGOF(Model3, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3, main = "GOF - Model 3 \n Indegree Distribution")
```

# Model 4

```{r, hide=TRUE}
Model4 <- fload("./data/processed/estimated_M4_20241123.rda")
```

```{r GOF M4, hide=TRUE}
GOF_M4 <- sienaGOF(Model4, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M4, main = "GOF - Model 4 \n Indegree Distribution")
```