### 2024-11-25 finalestimate.Rmd

---
title: "Model Estimation"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
editor_options: 
  markdown: 
    wrap: 72
---

<html lang="en">

<head>

<meta charset="UTF-8">

<meta name="viewport" content="width=device-width, initial-scale=1.0">

<title>My Models</title>

<link rel="stylesheet" href="styles.css">
<!-- Link to the CSS file I created for dark mode code chunks -->

</head>

</html>

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Setup

Clean workspace. As always.

```{r clean, echo=FALSE}
rm(list = ls())
```

## Packages

```{r package, include=TRUE, echo=FALSE, eval=TRUE}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

#install.packages("data.table")
require(data.table)  # mainly for faster data handling
# install.packages('xml2')
require(xml2)
# install.packages('rvest')
require(rvest)
require(igraph)
require(ggraph)

#install.packages('visNetwork')
require(visNetwork)
#install.packages('threejs')
require(threejs)
#install.packages('networkD3')
require(networkD3)

```

## Functions

```{r functions 1, echo=FALSE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology"+"political science", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

```{r functions 2, echo=FALSE}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

## Get Data

Load Scholar Data

```{r scholar data}
scholars <- fload("./data/processed/scholars_20240924.rda") 
```

Save the output of the function.

```{r saving output}
# save the output of your function
test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
```

Load Ego Data

```{r df_ego data, include= FALSE}
df_ego <- fload("./data/processed/df_ego_20241108.rda")

```

# Make Network Data

```{r RSiena network data}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

# put nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

```

------------------------------------------------------------------------

# Analysis - Model Estimation

## \<\> Model 0 - Null Model

First, I estimated a null model with only reciprocity and degree. The
maximum convergence ratio of the model is 0.084. The GOF on the Indegree
Distribution is zero**.** This indicates that the model is a very bad
fit for the data and estimates of this model do not give us reliable
information.\
\
*For all GOF plots compiled in one place, see here:
<https://treesdude.github.io/labjournal/finalGOF.html>.*

For the sake of practise, here they are anyways:\
The outdegree (density) effect assesses the tendency of nodes to
establish outgoing connections, serving as a baseline for the likelihood
that any given node will initiate an outgoing tie. In the context of
this study, these outgoing ties represent the initiation of
co-publications. The outdegree estimate is significant and negative (b =
-2.6388, SE = 0.1460), suggesting that a scholar's tendency to initiate
collaboration is low and dense networks are less sought after. The
reciprocity parameter id significant and positive (b = 1.8035, SE =
0.4343), which implies that directed ties are likely to be reciprocated.
Specific to our case, it describes that scholars tend to initiate
collaborations with those alters who sent a tie to them too.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m0}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M0 <- sienaDataCreate(nets)

```

Print short descriptive report

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r}
print01Report(mydata_M0, modelname = "./results/results_M0") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M0}
# define effects structure 
myeff_M0 <- getEffects(mydata_M0)

```

### ---Step 3: Add Effects

Null model: none.

### ---Step 4: Specify model/algorithm

```{r algo M0}

myAlgorithm_M0 <- sienaAlgorithmCreate (projname = "results_M0")

```

### ---Step 5: Estimate model

```{r estimate M0, eval=FALSE}
# estimate the model 
Model0 <- siena07(myAlgorithm_M0, data = mydata_M0, effects = myeff_M0, returnDeps = TRUE)
Model0

fsave(Model0, file ="estimated_M0")
```

```{r load M0 output, echo=FALSE}
Model0 <- fload("./data/processed/estimated_M0_20241123.rda")
Model0
```

```{r GOF M0}
GOF_M0 <- sienaGOF(Model0, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M0, main = 'GOF - "Null" Model 0 \n Indegree Distribution')
```

------------------------------------------------------------------------

## \<\> Model 1 - Structural Effects

This model additionally entails InPop, outAct, isolateNet to control for
effects of network statistics. The maximum convergence ratio of this
model is **0.17** and a Goodness of Fit on the indegree distribution is
**0.002.** Estimates of this model do not fit the data well and thus do
not give us reliable information.

For the sake of practise, here they are anyways:\
As in our null model, the parameters for outdegree (b = -2.0862, SE =
0.7050) and reciprocity (b = 1.0894, SE = 0.4144) remain significant and
unchanged on their 'direction'. The added parameter for indegree
popularity tells us something about the popularity of people who have a
lot of outgoing ties. In our model the estimate for this parameter is
significantly positive (b =0.2549, SE =0.0439), suggesting that scholars
specifically like to initiate collaborations with those who have a lot
of indegrees.

Further, the parameter for outdegree activity conveys information about
scholars' preferences to collaborate with others who have a lot of
outdegrees. In our case, the estimate is insignificant and cannot be
reliably interpreted. Lastly, the network isolate parameter can be used
to infer about scholars' tendency to neither send nor receive ties. The
estimate for this parameter is significantly positive (b = 4.9645, SE =
1.1146) in our model, indicating that the scholars of our sample tend to
remain isolates.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define M1}
# dependent variable 
nets <- sienaDependent(net_array)

# define data object 
mydata_M1 <- sienaDataCreate(nets)

```

```{r, echo=FALSE}
# make folder 
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r,echo=FALSE}
#Print short descriptive report
print01Report(mydata_M1, modelname = "./results/results_M1") 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m1}
# define effects structure 
myeff_M1 <- getEffects(mydata_M1)

```

### ---Step 3: Add Effects

```{r effects m1}

myeff_M1 <- includeEffects(myeff_M1, isolateNet, inPop, outAct)

```

### ---Step 4: Specify model/algorithm

```{r algo m1}

myAlgorithm_M1 <- sienaAlgorithmCreate (projname = "results_M1")

```

### ---Step 5: Estimate model

```{r estimate m1, eval = FALSE}
# estimate the model 
Model1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, returnDeps = TRUE)
Model1
# maximum convergence >.25
# run again:
Model1.1 <- siena07(myAlgorithm_M1, data = mydata_M1, effects = myeff_M1, prevAns= Model1, returnDeps = TRUE)
Model1.1

#fsave(Model1, file ="estimated_M1")
#fsave(Model1.1, file = "estimated_M1.1")

```

```{r load M1 output, echo=FALSE}

Model1 <- fload("./data/processed/estimated_M1_20241123.rda")
Model1

Model1.1 <- fload("./data/processed/estimated_M1.1_20241123.rda")
Model1.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M1}
GOF_M1.1 <- sienaGOF(Model1.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M1.1, main = "GOF - Model 1 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 2 - Gender

This model is partly similar to Model 1 but additionally holds the
network statistics sameX, diffXTransTrip, egoX & altX of the covariate
gender. EgoX and AltX ought to indicate whether male or female scholars
are more likely to initiate or receive ties, while sameX models the
general tendency for scholars to collaborate with same-gender peers. The
statistic for transitive triplets based on a different covariate value
(here: different gender) was originally included to control for
relationships of triadic nature (diffXTransTrip), but will be
disregarded in the interpretation of the results since a more
appropriate control statistic would have been that of transitive
triplets (TransTrip), but I simply lack the time to adjust the models.

The maximum convergence ratio for model 2 is **0.132** and the Goodness
of Fit on the indegree distribution is **0.096**. Looking at our
'base-line' network statistics, it is noteworthy that the estimate for
Outdegree is not significant anymore. This statistic's prior
significance can seemingly be explained by a parameter relating to the
covariate gender that I added to this model. However, the estimate of
the outdegree parameter is still nearing the significance 'threshold' (b
= -1.8628, SE = 1.1060, b/SE = -1.684268).

Comparing this model to the prior ones, we see that the reciprocity
estimate is still significantly positive (b = 1.3341, SE = 0.4543). The
parameters of the structural network statistics also show similar
patterns in estimates and significance as they did in model 1.
Concerning the added covariate "gender", we find that the parameter
estimates for both covariate-related activity (egoX) , pertaining to the
gender of the ego, and covariate-related popularity (altX), relating to
the gender of the alter, do **not** show any significance and hence do
not allow for reliable inferences. The 'same-covariate' measure (sameX)
for gender is significantly positive (b = 0.6637, SE =0.2685). It
reveals that scholars, indeed, prefer same-gender collaborators. This
supports H1, which posits gender homophily in collaboration patterns.

Model 2 provides strong evidence for gender homophily (H1), with
scholars more likely to collaborate with same-gender peers, but no
significant differences in collaboration tendencies based on the gender
of ego or alter.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m2}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

# define data object 
mydata_M2 <- sienaDataCreate(nets, gender_RS)

```

Print short descriptive report

```{r, echo=FALSE}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r, echo=FALSE}
print01Report(mydata_M2, modelname = "./results/results_M2") # create file with report in results folder 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m2}
# define effects structure 
myeff_M2 <- getEffects(mydata_M2)
```

### ---Step 3: Add Effects

```{r effects m2}
myeff_M2 <- includeEffects(myeff_M2, isolateNet, inPop, outAct)

myeff_M2 <- includeEffects(myeff_M2, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") 
```

### ---Step 4: Specify model/algorithm

```{r algo m2}

myAlgorithm_M2 <- sienaAlgorithmCreate (projname = "results_M2")

```

### ---Step 5: Estimate model

```{r estimate m2, eval =FALSE}
# estimate the model 
Model2 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, returnDeps = TRUE)
Model2
# maximum convergence >.25
# run again:
Model2.1 <- siena07(myAlgorithm_M2, data = mydata_M2, effects = myeff_M2, prevAns= Model2, returnDeps = TRUE)
Model2.1

#fsave(Model2, file ="estimated_M2")
#fsave(Model2.1, file = "estimated_M2.1")

```

```{r load M2 output, echo=FALSE}

Model2 <- fload("./data/processed/estimated_M2_20241123.rda")
Model2

Model2.1 <- fload ("./data/processed/estimated_M2.1_20241123.rda")
Model2.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M2}
GOF_M2.1 <- sienaGOF(Model2.1, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M2.1, main = "GOF - Model 2 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 3 - Department

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
investigate differences in same-gender collaboration patterns based on
the two departments.

Originally, I planned to also include an interaction term of departm_RS
and gender_RS to obtain clearer insights into variations in patterns of
gender homophily in relation to scholars' departments. However, I have
refrained from adding another model to this project. I decided against
this partly due to time constraints but I also reason that I can use the
comparison of model 2 and 3 to examine changes in the estimate of
'sameX' (gender), to *tentatively* infer whether the department of
scholars modifies the collaboration patterns at hand.

Model 3 has a maximum convergence ratio of **0.182** and a Goodness of
Fit on the indegree distribution of **0.108**.

Interestingly, the estimate for outdegree (density) that had lost its
significance in model 2, is now back to being significant and negative
(b = -4.3233, SE = 1.0871).

In terms of estimate size and significance of the structural network
statistics we, once again, find similar patterns as in models 1 and 2.
However, the outdegree parameter has now reached significance (b =
-4.3233, SE = 1.0871).

With regard to the network statistics of the covariate 'gender', we find
that the estimates relating to the gender of an ego (egoX) and the
gender of an alter (altX) remain non-significant. The parameter
estimates for sameX remains significantly positive (b = 0.6632, SE =
0.2623) and are almost identical to those of model 2 (b = 0.6632, SE =
0.2623). This suggests that adding the covariate 'department' did not
bring about any noteworthy changes in patterns of scholars seeking out
same-gendered scholars to collaborate.

When it comes to the parameters pertaining to the added covariate
'department', we find that any differences in collaboration based on the
department of an ego (egoX) or of an alter (altX) are not warranted, as
the estimates of these parameters are insignificant. However, we find a
significantly positive estimate for the parameter of 'covariate-related
identity' (sameX), which indicates that scholars are more likely to
collaborate with people from their own department, indicating
**department-based homophily**. (b = 2.7240, SE = 0.7277).

Model 3 highlights persistent gender homophily (H1) across departments
but does not provide evidence for differences in gender homophily
between Sociology and Political Science, contrary to H2.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m3}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender

# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# define data object 
mydata_M3 <- sienaDataCreate(nets, gender_RS, departm_RS)

```

```{r, echo=FALSE}
#create folder
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r, echo=FALSE}
#Print short descriptive report
print01Report(mydata_M3, modelname = "./results/results_M3") #
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff m3}
# define effects structure 
myeff_M3 <- getEffects(mydata_M3)
```

### ---Step 3: Add Effects

```{r effects m3}
myeff_M3 <- includeEffects(myeff_M3, isolateNet, inPop, outAct) # model 1

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M3 <- includeEffects(myeff_M3, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

```

### ---Step 4: Specify model/algorithm

```{r algo m3}

myAlgorithm_M3 <- sienaAlgorithmCreate (projname = "results_M3")

```

### ---Step 5: Estimate model

```{r estimate m3, eval =FALSE}
# estimate the model 
Model3 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, returnDeps = TRUE)
Model3

#if maximum convergence >.25: run again
Model3.1 <- siena07(myAlgorithm_M3, data = mydata_M3, effects = myeff_M3, prevAns= Model3, returnDeps = TRUE)
Model3.1

#fsave(Model3, file ="estimated_M3")
#fsave(Model3.1, file = "estimated_M3.1")

```

```{r load M3 output, echo=FALSE}

Model3 <- fload("./data/processed/estimated_M3_20241123.rda")
Model3

Model3.1 <- fload ("./data/processed/estimated_M3.1_20241123.rda")
Model3.1
```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M3}
GOF_M3.1 <- sienaGOF(Model3.1 , IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M3.1, main = "GOF - Model 3 \n Indegree Distribution")
```

------------------------------------------------------------------------

## \<\> Model 4 - H-Index

This model further builds on the prior model and also includes network
statistics of the covariate H-Index, namely egoX, altX and inPopX. By
adding these effects we can model how the h-index of a scholar
influences their preferences regarding same-gendered collaborations.
Similarly, to model 3, this model could have profited from a 'follow-up'
model that includes and interaction term between h-index and gender to
better infer about their influences. The same reasoning for my decision
to stick with one model prior also applies to this case. EgoX for this
covariate examines how the h-index of a scholar affects their
collaboration behavior, while AltX examines how the h-index of potential
collaborators affects their attractiveness as partners. With the help of
these effects we can test prestigious scholars are more likely to
initiate or receive ties. By looking at in-degree popularity weighted by
sender’s H-index (inPopX) we can assess whether a high h-index increases
the likelihood of being sought out as a collaborator.

Model 4 has a maximum convergence ratio of **0.23,** and a a Goodness of
Fit on the indegree distribution of **0.197**.

**Structural network statistics.** The estimate for outdegree is
negative but not significant. This suggests that, after accounting for
other variables, the baseline likelihood of forming ties remains low and
is not a major driver of collaboration patterns. Reciprocity has a
significant positive estimate (b = 2.1770, SE = 0.5960). This reflects a
strong tendency for mutual collaborations, meaning if one scholar
collaborates with another, the latter is likely to collaborate back
(dyadic reciprocity). The paramters Indegree Popularity and Outdegree
Activity are both insignificant, suggesting that neither the tendency to
receive many collaborations (indegree popularity) nor the tendency to
initiate them (outdegree activity) is significantly relevant to the
network structure. The network isolate estimate is, once again,
significantly positive ( b = 5.4634, SE = 2.2770), indicating that as
isolates tend to remain isolated.

**Gender.** Similar to AltX and EgoX in prior models, both an alter's
gender (b = -0.0903, SE = 0.3372) and an ego's gender (b = 0.2993, SE =
0.4934) still have insignificant effects. The parameter SameX shows a
significant positive estimate (b = 0.6820, SE = 0.2804) which barely
differs from the two prior models. This suggests that adding the
covariate 'H-Index' did not prompt any major changes in scholars'
preferences for same-gendered collaborators.

**Department.** While an alter's departmen**t** (AltX) remained
insignificant to the collaboration patterns, the estimates for an ego's
department changed incomparison to the prior model. In this model, egoX
seems to be of significant relevance (b = -1.4652, SE = 0.6963), with
its negative estimate indicating that, once effects of h-index are
controlled for, scholars from the political science (coded as 1)
department have a lower tendency to initiate collaborations than
scholars from the sociology department (coded as 0).\
Similar to the prior model, the sameX parameter for department shows a
significant positive estimate (b = 2.6327, SE = 0.6080) scholars are
still more likely to collaborate with others from the same department.

**H-Index**. Both an alter's and an ego's H-Index seem to be relevant in
collaboration preferences. The significant and positive estimateof AltX
(b = 0.0829, SE = 0.0364) suggest that scholars with higher h-index are
more preferred as collaborators. The egoX estimate is significant and
negative (b = -0.1591, SE = 0.0744) suggesting that scholars with higher
h-index are less likely to initiate collaborations.

The results partially support **H4**, as the h-index is shown to
influence collaboration patterns, with prestigious scholars being more
likely to attract collaborations as alters but less likely to initiate
them as egos. However, gender homophily remains significant even after
controlling for the h-index, suggesting that prestige does not
substantially weaken the influence of gender homophily in co-authorship
networks.

### ---Step 1: Define Siena Data & Variables

```{r RSiena define m4}
# dependent variable 
nets <- sienaDependent(net_array)

# put gender in object 
gender <- df_ego$gender
# covariate 1
gender_RS <- coCovar(gender) # gender as covariate in Rsiena

#put department in object 
department <- factor(df_ego$discipline.24, levels = c("sociology", "political science"))
#make numeric
department_numeric <- as.numeric(department) - 1 # sociology = 0, political science = 1
# covariate 2
departm_RS <- coCovar(department_numeric)

# put h-index in object 
h_index <- df_ego$h_index
df_ego$h_index
# covariate 3
h_index_RS <- coCovar(h_index) # h_index as covariate in RSiena

# define data object 
mydata_M4 <- sienaDataCreate(nets, gender_RS, departm_RS, h_index_RS)

```

```{r, echo=FALSE}
#create folder
ifelse(!dir.exists("results"), dir.create("results"), FALSE)

```

```{r, echo=FALSE}
#Print short descriptive report
print01Report(mydata_M4, modelname = "./results/results_M4") 
```

### ---Step 2: Make effects object/structure

```{r RSiena myeff M4}

# define effects structure 
myeff_M4 <- getEffects(mydata_M4)

```

### ---Step 3: Add Effects

```{r effects M4}

myeff_M4 <- includeEffects(myeff_M4, isolateNet, inPop, outAct) # model 1

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, diffXTransTrip, sameX, interaction1 = "gender_RS") #model 2

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, sameX, interaction1 = "departm_RS") #model 3

myeff_M4 <- includeEffects(myeff_M4, egoX, altX, inPopX, interaction1 = "h_index_RS") # model 4

```

### ---Step 4: Specify model/algorithm

```{r algo M4}

myAlgorithm_M4 <- sienaAlgorithmCreate (projname = "results_M4")

```

### ---Step 5: Estimate model

```{r estimate M4, eval =FALSE}
# estimate the model 
Model4 <- siena07(myAlgorithm_M4, data = mydata_M4, effects = myeff_M4, returnDeps = TRUE)
Model4

#fsave(Model4, file ="estimated_M4")

```

```{r load M4 output, echo=FALSE}

Model4 <- fload("./data/processed/estimated_M4_20241123.rda")
Model4

```

### ---Step 6: GOF - Goodness of Fit

```{r GOF M4}
GOF_M4 <- sienaGOF(Model4, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "nets")

plot(GOF_M4, main = "GOF - Model 4 \n Indegree Distribution")
```