### 2024-11-25 index.Rmd
finalpushh

---
title: "Lab journal"
author: "by: Vero Palm"
bibliography: references.bib
---

```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
# install.packages("remotes")
# remotes::install_github("rlesur/klippy") # is what makes it possible to hover over top right of code on website to COPY AND PASTE it 
```

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r token, include=FALSE}
#install.packages("gitcreds")
# gitcreds::gitcreds_set()
```

::: columns
::: {.column width="50%"}
# First page of something totally awesome yes.
Last compiled on `r format(Sys.time(), '%B, %Y')`
*Description: Very cool lab journal. Yes, yes.*

## Cool keybinds

[CTRL + ALT + I to make new chunks]{.smallcaps}

[ALT + MINUS makes \<-]{.smallcaps}

[STR + SHIFT + M makes the %\>%]{.smallcaps}

[STR + SHIFT + C can make a comment of multiple lines you highlighted]{.smallcaps}

```{=html}
<!--- The ---- is a visual line (page break looking mf) on the 
website 
Always ADD A SPACE BETWEEN A PAGE BREAK AND the next comment, otherwise it will appear on page --->
```
```{=html}
<!--- This is the best way to write a comment. 
It also demonstrates, you can add html directly in your rmarkdown file. See: https://www.w3schools.com/html/html_comments.asp --->
```
```{=html}
<!--- to use a multi column page we use Pandoc's fenced Div. 
see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html 
But please note, we are not here to make a beautiful website, our intention is just to have a quick lab journal up and running --->
```

## Open Science

Everyone can fork this repository. I have used this example to journal my lab work and assignments.

Jochem Tolsma and Bas Hofstra use this repository for their Social Network Analysis For Social Scientists course [@SNASS]. <!--- this is a reference, stored in references.bib. It will be automatically added at the bottom of the page --->

:::

::: {.column width="50%"}
![](images/krill-issue.jpg)

Very krillaxing, isn't it?

:::
:::

**Reference:**